### 2015-12-06 Parallel CPU Implementation of Mesh Free Method on 3D Multiphase Mixture Flow .tex
my first commit

\documentclass[10pt,a4paper]{article}
%\documentclass[10pt,a4paper]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{kpfonts}
\usepackage[justification=centering]{caption} %to adjus caption center aligned
\usepackage{hhline} %to add double horizontal line in tabular
\usepackage[numbers]{natbib}
%other options include: round, curly, angle, or default options
\usepackage{media9} %package for include movies
\author{Zhixuan Cao, Abani Patra}
\title{Data Management and Domain Adjusting Strategies for Implementing Parallel SPH Method in JPUE Simulation}

\begin{document}

\maketitle
%\tableofcontents
\section{Abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and decompose (redecompose) the domain. Hilbert Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC indexes are assigned to particles to guarantee uniqueness of identifier.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are stored in hashtables. Additional link list attached to the hashtable is used to handle hash confliction. A dynamic domain decomposition strategy is implemented to ensure good load balance. Furtherly, computational domain is adjusted during simulation to reduce computational cost. 
Numerical tests show that our code has good strong scalability.
These strategies can be further applied to many other implementations of meshfree methods, Especially those implemetations that require flexible accessing, adding and deleting of particles.
\section{Introduction}
SPH is a meshless scheme, initially developed for astrophysical applications by Lucy\citep{lucy1977numerical} and Gingold and Monaghan\citep{gingold1977smoothed}. 
Subsequently it was extended to large strain solid mechanics and computational fluid mechanics. 
A lot of research has been done in parallellization of SPH. 
Goozee\citep{goozee2003distributed} implemented a simple SPH code using MPI, OpenMP and BSP. 
Chen Wenbo\citep{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
Jose M. David W. Holmes\citep{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
Domínguez\citep{dominguez2011optimization} present optimizations for both CPU and GPU parallelization of a SPH method. 
Angela Ferrari\citep{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme. 
D. Kumar and Abani\citep{kumar2013parallel} implemented CPU parallel Godunov SPH in simulation of granular flows.
A.J.C. Crespo\citep{crespo2015dualsphysics} used the parallel power computing of GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
However, most implementation of parallel SPH method is limitted to bechmark problems like dam break, or relative simple scenarios like breaking-waves, floodings ect. Rare work has been done in simulation of more complicated problems, such as eruption of volcano plume which is essentially a mutilple phases, turbulent, ejection flow process accompanied with microphysics phenomena like phase change of water, aggregation, reaction ect. Prediction of such more complicated phenomena with acceptable accuracy at given time window can not been acomplished without parallel computing. Imposing of boundary conditions (such as realistic wind feild, eruption boundary condition and no-slip wall boundary condition) requires dynamiclly adding and removing of particles during simulation. This requies efficient and more flexible data management scheme. In this paper, we implement SPH to simulate a simplified version of this complicated volcanic plume: JPUE. and develop data management methodologies for it.\\
Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing.  (eg.\citep{ferrari2009new, crespo2015dualsphysics}). Less attention has been paid to developing of more flexible data structures for more complicated problem. 
Fortunately, efficient and flexible data management strategies for high performance computing have been successfully implemented in meshbased methods(eg. \citep{laszloffy2000simple} for adaptive hp FEM, and \citep{pitman2003computing, patra2005parallel} for FVM). Motivated by techniques developed for mesh based methods, we present a complete framwork for parallelizing SPH program with MPI standard model allowing more flexible and efficient data access in this paper.\\
Any implementation of SPH code requires efficient searching and updating of neighbors during simulation. We adopted background grid which was proposed by Monaghan and Lattanzio\citep{monaghan1985refined} and implemented widespreadly in parallel SPH. The background grid is also used for domain decomposition in SPH. We refer to the elements of background grid, namely a squares for two dimension and cubics for three dimension, as buckets. As for the storage of physical quantities associated to each particle, different strategies was adopted in existing implementation of SPH. 
In both SPHysics\citep{dominguez2011optimization} and DualSPHysics\citep{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells. This has two advantage: 1)access pattern is more regular and more efficient, 2) It is ease to identify the particles that belongs to a cell by using a range since the first particle of each cell is known. But adding, deleting and especially accessing of particles are cumbersome. Angela Ferrari\citep{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation. Storage problems caused by fix-size arrays are thereby also eliminated. We define C++ classes which contains all data of particle and bucket. As for the storage of data, we adopt an hashtable to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list. Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier. The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have a different identifier. As for domain decomposion, even though more complicated graph-based partition tools\citep{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming. So we adopt an easy-programming scheme based on SFC \citep{patra1999efficient}.\\
To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluid. This feature is accomplished by adding a scan function to monitor the most outside layer of the domain and turn ghost particles to real particles at proper time. 
%The computational cost of this additional scan function is low.
\\
The data structure, particle and bucket indexing strategies, domain decomposition and adjusting strategies, and dynamic load balancing method in this paper can be easily adopted by other implementations of any meshfree methods(include SPH). The flexibility of data accessing enables implementing of meshless methods for solving of more complicated problems.\\
%In the following sections, the physics and SPH implementation of inject flow problem is given, then 
\section{Data structure and load balance}

\begin{figure}[h]
\caption{Basic Work flow for SPH}
\centering
\label{fig:Work_flow}
\includegraphics[width=0.5\textwidth]{Work_flow}
\end{figure}

SPH is a meshfree, Lagrangian method. The domain is discretized by particles and the position of particle is updated at every time step. The physical laws (such as conservation laws of continuum fluid dynamics) written in the form of partial differential equations need to be transformed into the Lagrangian formalism. Using a kernel function that provides the weighted estimate of the field variables at a discrete point (particle), the integral equations are evaluated as sums over neighbor particles. Only particles located within support of kernel function will interact. Thus, physical proterties (position, density, velocity, internal energy, pressure) are updated based on its neighbors. So a neighbor searching need to be carried out before updating of physical proterties. We use buckets which overlap the whole domain and keep fixed in time during the entire simulation, to reduce searching cost. Domain decomposition will be based on SFC going through centroid of buckets. A basic works flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\

Particles need to be added and(or) removed during simulation of JPUE problem. Wall boundary conditions in SPH can be imposed either by adding a force term or using ghost particles. We adopt the latter one in our simulation. As our computational domain will be adjusted during simulation, wall ghost particles need to be added or deleted during simulation. New particles also need to be added for the eruption boundary condition. 
We will describe in this section these strategies which allows flexible deleting and adding.
\subsection{SFC based indexing}
Our data structure start from assigning each particle and bucket an identifier, we refer to it as key, which should be unique throughout simulation. The key for bucket is determined by centroid coordinates of the bucket while the key for particle is determined by adding coordinates and adding time step. The map from coordinates to key is based on SFC.\\
The SFC \citep{sagan2012space} maps n-dimentional space to a one dimentional sequence. The standard procedure for obtainning SFC is: 
\begin{itemize}
\item Scale coordinates of particle or bucket into $[0,1]^n $ based on maximum and minimum coordinates of the computational domain: $\textbf{X}^\prime \rightarrow \textbf{X}$
\item Computed $k_r = h_n(\textbf{X})$. Where $h_n$ is the map $h_n: [0,1]^n \rightarrow [0,1]$. 
\item Convert $k_r$ to integer $k$ by multiplying $k_r$ with a very large number and removing decimal part.
\item All keys are sorted to form a sequence which is SFC. SFC represents a curve passing through all particles (or centroid of buckets).

\end{itemize}
{
\begin{figure}[h]
\caption{Spcace filling curve orderings of buckets and particles in the buckets}
\centering
\label{fig:SFC_particles_buckets}
\includegraphics[width=0.6\textwidth]{SFC_particles_buckets}
\end{figure}

\begin{figure}[h]
\caption{Decomposition of domain based on SFC}
\centering
\label{fig:SFC_particles_buckets_partition}
\includegraphics[width=0.6\textwidth]{SFC_particles_buckets_partition}
\end{figure}
}

Scheme for constructing the map $h_n$ can be found in \citep{patra1995problem}. These keys denote a simple addressing/ordering scheme for the data and computations, i.e., a simple global index space for all the objects.\\
SFC-based indexing scheme can guarantee uniqueness of particle identifier only in simple scenarios when particles are added once while setting up initial condition. In some situations, new particles need to be added while simulation. For example, new particles need to be added at the bottom of the eject vent to simulate JPUE. To distinguish particles added at the "same place" (the small area, all points whithin which will be mapped to the same $k$.) from different time steps, we extend the SFC-based key to SFC based time-dependent key by including date of birth of particles into the key. The SFC based time-dependent key can be written as: $[k,t]$, where $t$ is the time step. The map $h_n$  will become: 
$h_n: [0,1]^n \times \textbf{T} \rightarrow [0,1] \times \textbf{N}$
where $\textbf{T} \subset [0,\infty)$ is the time step dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.
To guarantee locality, sorting of particle keys is majorly based on $k$, that is to say, particle with smaller $k$ always comes before particles with larger $k$. For these particles has the same $k$, ordering of them will depend on $t$. Figure \ref{fig:SFC_particles_buckets} shows SFC ordering of buckets and particles in buckets. 
Several features of such indexing scheme are suitable for SPH:
\begin{itemize}
\item Guarantee uniqueness of identifier.
\item Key of each object is generated purely based on its own coordinates. While add new objects on different processes, key of each object can be generated fast and independently .
\item Objects that locate closely in the Euclidian space will also be close to each other in the one dimensional key space in the mean sense. Since SPH particles only interact with its neighbors, geometric locality can be exploited for efficient storage and retrieval of bucket and particle data.
\item This type of key effectivelly generate a global address space. Globality of key and conservation of locality make it easy to partition the sorted key squence and obtain the decompositiion of the problem.
\end{itemize}
What need to point out is that motion of particles might mess up locality that established based on initial coordinates of particles. As particles are moving regularly, the locality of most of particles should still be conserved during simulation.
%If we can design a re-numberring method which is computationally cheap, we can restore the locality.
\subsection{Data structure}
\subsubsection{Particle and bucket}
The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbor searching and domain decomposition. Both are defined as classes in C++. Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(process that the particle belong to), primitive variables (variables show up as unknows in governing equations, like coordinates, density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg. pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is vector of particle keys in our application). Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (27 neighbor buckets including itself for three dimension and 9 for two dimension). Objects defined based on these two classes are then accesse through hash tables.
\subsubsection{Hash table and hash confliction}
As discussed at the beginning of this section, implementation of SPH in more realistic scenarios requires dynamic memory management and flexible data access. One of the fundemental data structures that satisfy such requirement is hash table. Another option is B-tree. We adopt hash table. An implementation of B-tree under a similar situation for mesh based methods can be found in other papers(eg.\citep{patra2003data}).\\
Hash table, which is divided into slots, are array based data structure. Based on the key, the address-calculator(hashing) function determines in which slot the data should be stored. The hashing function maps from key to the slot index:
\begin{equation}
slot\,index = hash(key)
\end{equation}
The hash table has O(1) data accessing, adding and delete properties when there is no confliction. How many conflictions will happen depends on both distribution of keys in the keys space and size of the hash table. As the distribution of keys is determined by particles' initial location (or centroids of buckets), the hash table size is under our control. We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to low prefetch and low memory efficiency. Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having many hash confliction. Abani\citep{patra2003data} did numerical experiments to make trade-off to determine hash table size.\\
 %We should also do some experiment to optimize our hash table size.
One way to handle hash confliction is using an additional sorted vector attached to the hash table. When several keys hash to the same slot, a vector will be created. The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrievaling . Another option to handle hash confliction is using an additional link list which is more flexible in memory allocation. The average time complexity of binary search is O(log n) while that for linear search based on link list is O(n). However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longger. Choosing of proper way to handle hash confliction greatly depends on the problem itself. For the test problem in this paper, successively adding of particles at the bottom of the eject vent will lead to hash confliction of many particles, which implies a long link list. But considering the very long conflictions only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle the hash confliction. This decision was made based on numerical experiments.
\subsubsection{Hash function}
\begin{figure}[h]
\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, these new particles will be stored in external link list}
\centering
\label{fig:Particle_adding_with_link}
\includegraphics[width=0.6\textwidth]{Particle_adding_with_link}
\end{figure} 
For time-independent keys, the hash function can be a simple function like:
\begin{equation}
\label{eq:hash_function}
slot\,index = \frac{Key - Min\,Key}{Max\,Key - Min\,Key} \times Hash\,Table\,Size 
\end{equation}
One natural way to hash SFC based time-dependent key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as higher digit and $t$ as lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent. That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as show in figure \ref{fig:Particle_adding_with_link}. To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number $k$ of the key $[k,t]$ into the hashing function, equation (\ref{eq:hash_function}).

\subsection{Load balance strategy}
\subsubsection{Weighted work load}
Particles in the test problem can be categorized into real particle, wall ghost particle, pressure ghost particle, eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}). As different type of particles involves different amount of computation,shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different type of particles based on profilling data. Work load of each bucket is then determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now become a weighted sequence. Domain decomposition will conducted based the weighted SFC.
\begin{table}[h!]
  \centering
  \caption{Computational cost for different steps per particle}
  \label{tab:Computational_cost_steps}
  \begin{tabular}{|c|r|r|}
    \hline
    Step & Cost ($\mu s$) & Abbreviation\\
    \hhline{|=|=|=|}
    neighbor search & 0.45 & NS\\
    \hline
    update momentum and energy & 0.62 & UPME\\
    \hline
    update density & 0.37 & UPD\\
    \hline
    update position & 0.02 & UPP\\
    \hline
    velocity filtering& 0.38 & VF\\
    \hline
    apply wall bc & 1.2 & WBC\\
    \hline
  \end{tabular}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Computational work load for each type of particle}
  \label{tab:Computational_cost_particles}
  \begin{tabular}{|c|r|r|r|r|r|r|r|}
    \hline
    Particle type & NS & UPME & UPD & UPP &VF &WBC &Total\\
    \hhline{|=|=|=|=|=|=|=|=|}
    Real & Yes & Yes & Yes & Yes & Yes & No & 1.8\\
    \hline
    wall ghost & No & No & No & No & No & Yes &1.2\\
    \hline
    eruption ghost & No & No & No & Yes & No & No & 0.02\\
    \hline
    pressure ghost & No & No & No & No & No & No & 0.\\
    \hline
  \end{tabular}
\end{table}
\subsubsection{Domain decomposition and dynamic load balancing}
Domain decomposition will be conducted based the weighted SFC of buckets. Figure \ref{fig:SFC_particles_buckets} and figure \ref{fig:SFC_particles_buckets_partition} show how domain will be decomposed based on partition of SFC of buckets. The particles are automatically split into two groups along with buckets that contain them. In our current implementation, the domain is partitioned only in $x-y$ plane. A two dimensional footprint of the three dimensional domain(as well as buckets) is generated by projecting. 
%SFC for the two dimensional footprint is partitioned for domain decomposition. 
The work load of each object on the SFC of footprint buckets in $x-y$ plane will be the summation of work loads of all buckets above the footprint. 
Movement of particles, adding of new particles, adjusting of domain will lead to load imbalance between processes. Computational load is monitored at given interval (The interval is optimized based on numerical experiments). Repartitioning is carried out if load imbalance is larger than a given tolenrance. Figure \ref{fig:t0_domain_decomposition} and \ref{fig:t10_domain_decomposition} show the initial decomposition and domain decomposion at 10 seconds. 
%We adopt the same SFC based partition strategy as described by Patra \ref{patra1999efficient}. 
\begin{figure}[h]
\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: initial domain decomposition}
\centering
\label{fig:t0_domain_decomposition}
\includegraphics[width=0.5\textwidth]{t0_domain_decomposition}
\end{figure}

\begin{figure}[h]
\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: decomposition at t=10s}
\centering
\label{fig:t10_domain_decomposition}
\includegraphics[width=0.5\textwidth]{t10_domain_decomposition}
\end{figure}
As some of the neighbor particles reside in other
partitions. A set of “guest” particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 
\section{Adjusting of domain during simulation}
%SPH particles can be viewed as discretization points\citep{diltsg1999moving}. 
As a Lagrangian method, SPH is able to automatically adjust computational domain(eg, dam break simulation) as the postition of the dicretization points are updated at every time step. However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone. Because the whole domain, which occupied by stationary fluid before eject fluids reaching there, has to be discretized at the very beginning of simulation. A lot of CPU time will be spent on computing of "stationary" particles. It is pure wasting of computational resources. If we can avoid simulating of stationary particls by adjusting the computational domain, the computational cost will be reduced greatly. To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. We proposed a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain,  ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature. (see figure \ref{fig:Work_flow_adjust})

\begin{figure}[h]
\caption{Work flow that enables domain adjusting feature}
\centering
\label{fig:Work_flow_adjust}
\includegraphics[width=0.5\textwidth]{Work_flow_adjust}
\end{figure}
The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class. Particles are categorized into three groups based on the value of the flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0). The involved particles are particles that have already been involved in eject mixing. The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles. So they will be involved in the mixing process in the near future. These not involved particles are particles that far away from the eject source and is still at the initial state. For these not involved particles, it is not necessary to update its state variables. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary. That is to say, only potential involved and involved particles need to be simulated.
As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced. The criteria to determine whether a particle is involved or not is: whether the mass fraction of that particle is larger than a given threshold (10e-5 in out simulation). Other state variable, such velocity, can also serve as alternative "swich criteria".\\

{
\begin{figure}[h]
\caption{Imposing of all kinds of boundary condition, and using of flag Involved}
\centering
\label{fig:Involved_initial}
\includegraphics[width=0.5\textwidth]{Involved_initial}
\end{figure}

\begin{figure}[h]
\caption{Adjusting of domain}
\centering
\label{fig:Involved_domain_adjusting}
\includegraphics[width=0.5\textwidth]{Involved_domain_adjusting}
\end{figure}
}

A has-involved-flag is added in bucket class. Buckets are also categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And all of its neighbor buckets will be set to be has potential involved. 
All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles. There are two situations that will make a has potential involved bucket to be a has involved bucket. First, when a involved particle enters a has potential involved bucket. Second, when mass fraction of a particle (It should be a potental involved particle) exceeds the threshold, this particle will turn to a involved particles and if this is the first involved particle in the bucket, the bucket will turn from a has potential involved bucket to a has involved bucekt.\\
The most outside buckets layer of has-potential-involved buckets will be scaned every time after updating. If any a of these buckets becomes has involved, the domain will be enlarged by turnning the pressure ghost particles to real particles and turning involved-flag from 0 to 1. Also, the buckets that originally contain pressure ghost particles will become has potential involved. New pressure ghost particles and wall ghost particles will be added around the adjusted domain.
This process of domain adjusting is show in figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}.
The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}

For the test problem in this paper, the volcanic plume will finally reach to a domain of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption. When numerical simulation goes up to 90 seconds, the plume is still within a domain of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$. Adjusting of domain can avoid computing large number of uninfluenced air particles. Numercial test shows that simulation time is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
\section{Numerical test}
As we are targeting at developing data management and paralell strategies for more complicated implementation of SPH which demmands quick and flexible data access, delete and add. The test problem should have such demand. Volcanic eruption which is essentially a mutilple phases, turbulent, ejection mixing flow accompanied with microphysics processes requires more flexibile data management. We adopt a two phase volcanic plume model\citep{suzuki2005numerical} as our test problem. One phase is air while another phase is ejected material.
\subsection{Governing equations\citep{suzuki2005numerical} and boundary conditions}
Based on Navier-Stokes equations, the following assumptions are made:
\begin{itemize}
\item Molecular Viscosity is neglected since eddy viscosity due to turbulence is dominant. As heat conduction is much smaller than turbulent heat exchange \citep{oberhuber1998volcanic}, We only consider turbulent heat exchange.
\item Erupted material consist of solid with different size and mixture of gas (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2). Air (also a mixture of different gas) is assumed to be another phase (phase 1).
\item We assume immediate thermodynamic equilibrium so that no separate energy equation is needed for each phase. As a result, there is only one energy equation for both phases and heat transfer effect between different phases will be ignored.
\item We assume immediate dynamic equilibrium so that no separate momentum equation is needed for each phase. As a result, there is only one vector moment equation for both phases. Drag force term will not show up in the momentum equation as two phases always move with the same velocity.
\item Because of above assumptions, all other micro-physics process (like phase change of $H_2O$, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model.
\end{itemize}
Finally the governing equations for this model in Eulerian form are:
\begin{align}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho}\\
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks} \\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{align}
$\xi$ is the mass fraction of ejected material.
$E = e + K $ is total energy which is summation of kinetic energy $K$ and internal energy $e$.
An additional equation is required to close the system. In this model, the equation for closing the system is an EOS:
\begin{align}
p = (\gamma_m - 1)\rho e \label{eq:EOS}\\
\end{align}
Where 
\begin{align}
\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}\\
Rm = n_gR_g + n_aR_a  \label{eq:gov-Rm}\\
C_{vm} = n_s C_{vs} + n_g C_{vg} + n_a C_{va} \label{eq:gov-Cvm}\\
n_a = 1 - \xi \label{eq:gov-na}\\
n_g = \xi n_{g0} \label{eq:gov-ng}\\
n_s = \xi - n_g \label{eq:gov-ns}
\end{align}
Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscription $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\

In current model the initial domain is a 3D box. The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
At the vent, $T$, $\textbf{v}=\{0,0,150\}^T$, $p=1.01\times10^5 Pa$, $n_{g0}=0.05$ and mass discharge rate $\dot M$ is given. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the bounday. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, boundary condtions for density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Involved_initial} and figure \ref{fig:Involved_domain_adjusting}). 
\subsection{Discretized governing equations with SPH}
There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive view over SPH. We will not cover these basic theory of SPH in this paper.
The discretized governing equations with SPH are:
\begin{align}
<\rho_a^a>=\sum m_b w_{ab} (h_a) \label{eq:gov-sph-d1}\\
<\rho_i^{sg}>=\sum_j m_j w_{ij} (h_i) \label{eq:gov-sph-d2}
\end{align}
\begin{equation}
<\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+\textbf{g} \label{eq:gov-sph-v}
\end{equation}

\begin{equation}
<\dfrac{d e_{\alpha}}{d t}>=
 0.5\sum_b [m_b \textbf{v}_{\alpha b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
 +0.5\sum_j [m_j \textbf{v}_{\alpha b}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})] \label{eq:gov-sph-e}
\end{equation}
Where
$\rho_a^a$ is density of phase 1 (air).
$\rho_i^{sg}$ is density of phase 2 (erupted material).
$\rho=\rho^a + \rho^{sg}$ is density of mixture of phase 1 and phase 2.
\begin{align}
\textbf{v}_{\alpha b}=\textbf{v}_{\alpha}-\textbf{v}_b \\
\textbf{v}_{\alpha j}=\textbf{v}_{\alpha}-\textbf{v}_j
\end{align}
$w_{ab} (h_a)= w(\textbf{r}_a-\textbf{r}_b,h_a)$ is smoothing kernel.
%$\bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})$-->need double check
$\Pi$ is artificial viscosity term \citep{monaghan1992smoothed}.
Index $a$, $b$ is for phase 1.
Index $i$, $j$ is for phase 2.
Index $\alpha$, $\beta$ can be index of either phase 1 or phase 2.
The position of each particle is updated according to the following equation.
\begin{align}
\dfrac{d \textbf{r}_a}{dt} = \textbf{v} \label{eq:gov-update-pos}
\end{align}
Free surface flow are in nature turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result to a filtered velocity for position updating and additional turbulent terms in momentum and energy eqution.
\subsection{Solver performance}
Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0, 4000]$, with smoothing length of 200m. The computation was run for 20s physical time. The running time and speed up are shows in figure \ref{fig:performance_128_time} and \ref{fig:performance_128_speed_up}. The results shows very good strong scalability when number of processes is smaller than 64. A overhead of scalability also shows up in the figure. The over head of speed up is due to 2D domain decomposition. For this test case, the total number of footprint buckets in x-y plane is 64. Recall that domain decomposition 2D and is based on SFC of footprint buckets. When we request 128 processes but there is only 64 buckets initially, another 128 processes will be totally idle at the beginning. As simulation goes, the domain is expanded and more and more idle processes will start getting tasks. That's why we still get some speed up. This can be easily demonstrated by another experiment with more buckets on x-y plane. In our second experiment, we still use the same domain, but we take smoothing length as 100m. As a reseult, there will be 256 footprint bucekts on the x-y plane initially. The results are consistent with our expectation: we got linear speed up up to 256 processes. Adgain, the an overhead shows up when number of processes is larger than 256. This motivate us to adopt three dimensional domain decomposition in our future research.
\begin{figure}[h]
\caption{Running time Vs number of processes}
\centering
\label{fig:performance_128_time}
\includegraphics[width=0.7\textwidth]{128_time}
\end{figure}
\begin{figure}[h]
\caption{Speed up: Strong scalability}
\centering
\label{fig:performance_128_speed_up}
\includegraphics[width=0.7\textwidth]{128_speed_up}
\end{figure}

\section{Conclusion}
We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology that is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particles. 
Hashtables with external link list are adopted for accessing  particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The code was further improved to 4 times faster by adjusting computational domain according to progress of simulation. 
Our code shows a linear speed up and an overhead cast by number of footprint buckets in x-y plane. This motivate us to adopt three dimensional domain decomposition in our future research.

\bibliographystyle{plainnat}
\bibliography{Reference}

\end{document}

### 2015-12-06 Reference.tex
my first commit

%boundary conditions imposing for Hyperbolic system of equaiton.
@article{thompson1990time,
  title={Time-dependent boundary conditions for hyperbolic systems, II},
  author={Thompson, Kevin W},
  journal={Journal of Computational Physics},
  volume={89},
  number={2},
  pages={439--461},
  year={1990},
  publisher={Elsevier}
}
@book{godlewski2013numerical,
  title={Numerical approximation of hyperbolic systems of conservation laws},
  author={Godlewski, Edwige and Raviart, Pierre-Arnaud},
  volume={118},
  year={2013},
  publisher={Springer Science \& Business Media}
}
@article{kreiss1970initial,
  title={Initial boundary value problems for hyperbolic systems},
  author={Kreiss, Heinz-Otto},
  journal={Communications on Pure and Applied Mathematics},
  volume={23},
  number={3},
  pages={277--298},
  year={1970},
  publisher={Wiley Online Library}
}
@article{majda1975initial,
  title={Initial-boundary value problems for hyperbolic equations with uniformly characteristic boundary},
  author={Majda, Andrew and Osher, Stanley},
  journal={Communications on Pure and Applied Mathematics},
  volume={28},
  number={5},
  pages={607--675},
  year={1975},
  publisher={Wiley Online Library}
}
@article{ralston1971note,
  title={Note on a paper of Kreiss},
  author={Ralston, James V},
  journal={Communications on Pure and Applied Mathematics},
  volume={24},
  number={6},
  pages={759--762},
  year={1971},
  publisher={Wiley Online Library}
}
@article{benabdallah1987problemes,
  title={Probl{\`e}mes aux limites pour des syst{\`e}mes hyperboliques non lin{\'e}aires de deux {\'e}quations {\`a} une dimension d'espace},
  author={Benabdallah, Assia and Serre, Denis and LIONS, J-L},
  journal={Comptes rendus de l'Acad{\'e}mie des sciences. S{\'e}rie 1, Math{\'e}matique},
  volume={305},
  number={15},
  pages={677--680},
  year={1987},
  publisher={Elsevier}
}
@book{dubois1989boundary,
  title={Boundary conditions for nonlinear hyperbolic systems of conservation laws},
  author={Dubois, Fran{\c{c}}ois and Le Floch, Philippe},
  year={1989},
  publisher={Springer}
}
@techreport{oliger1976theoretical,
  title={Theoretical and Practical Aspects of Some Initial-Boundary Value Problems in Fluid Dynamics.},
  author={Oliger, Joseph and Sundstrom, Arne},
  year={1976},
  institution={DTIC Document}
}
%The paper that I started from to get a basic idea on  IBCP.
@article{guaily2013boundary,
  title={Boundary conditions for hyperbolic systems of partial differentials equations},
  author={Guaily, Amr G and Epstein, Marcelo},
  journal={Journal of advanced research},
  volume={4},
  number={4},
  pages={321--329},
  year={2013},
  publisher={Elsevier}
}
%Books on CFD
@book{pletcher2012computational,
  title={Computational fluid mechanics and heat transfer},
  author={Pletcher, Richard H and Tannehill, John C and Anderson, Dale},
  year={2012},
  publisher={CRC Press}
}

%Very old paper on volcano plume---> useful as in review.
@inproceedings{morton1956turbulent,
  title={Turbulent gravitational convection from maintained and instantaneous sources},
  author={Morton, BR and Taylor, Geoffrey and Turner, JS},
  booktitle={Proceedings of the Royal Society of London A: Mathematical, Physical and Engineering Sciences},
  volume={234},
  number={1196},
  pages={1--23},
  year={1956},
  organization={The Royal Society}
}
@book{bullard1962volcanoes,
  title={Volcanoes: in history, in theory, in eruption},
  author={Bullard, Fred Mason},
  year={1962},
  publisher={Wiley Online Library}
}
@article{settle1978volcanic,
  title={Volcanic eruption clouds and the thermal power output of explosive eruptions},
  author={Settle, Mark},
  journal={Journal of Volcanology and Geothermal Research},
  volume={3},
  number={3},
  pages={309--324},
  year={1978},
  publisher={Elsevier}
}
@article{wilson1978control,
  title={The control of volcanic column heights by eruption energetics and dynamics},
  author={Wilson, L and Sparks, RSJ and Huang, TC and Watkins, ND},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={83},
  number={B4},
  pages={1829--1836},
  year={1978},
  publisher={Wiley Online Library}
}
@article{woods1988fluid,
  title={The fluid dynamics and thermodynamics of eruption columns},
  author={Woods, AW},
  journal={Bulletin of Volcanology},
  volume={50},
  number={3},
  pages={169--193},
  year={1988},
  publisher={Springer}
}
@article{woods1995decompression,
  title={The decompression of volcanic jets in a crater during explosive volcanic eruptions},
  author={Woods, Andrew W and Bower, Sally M},
  journal={Earth and Planetary Science Letters},
  volume={131},
  number={3},
  pages={189--205},
  year={1995},
  publisher={Elsevier}
}
@book{sparks1997volcanic,
  title={Volcanic plumes},
  author={Sparks, Robert Stephen John and Bursik, MI and Carey, SN and Gilbert, Jennifer and Glaze, LS and Sigurdsson, H and Woods, AW},
  year={1997},
  publisher={Wiley}
}
@article{woods1993moist,
  title={Moist convection and the injection of volcanic ash into the atmosphere},
  author={Woods, Andrew W},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={98},
  number={B10},
  pages={17627--17636},
  year={1993},
  publisher={Wiley Online Library}
}
@article{koyaguchi1996formation,
  title={On the formation of eruption columns following explosive mixing of magma and surface-water},
  author={Koyaguchi, Takehiro and Woods, Andrew W},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={101},
  number={B3},
  pages={5561--5574},
  year={1996},
  publisher={Wiley Online Library}
}
@article{woods1991particle,
  title={Particle fallout, thermal disequilibrium and volcanic plumes},
  author={Woods, Andrew W and Bursik, Marcus I},
  journal={Bulletin of Volcanology},
  volume={53},
  number={7},
  pages={559--570},
  year={1991},
  publisher={Springer}
}

@article{valentine1989numerical,
  title={Numerical models of Plinian eruption columns and pyroclastic flows},
  author={Valentine, Greg A and Wohletz, Kenneth H},
  journal={J. Geophys. Res},
  volume={94},
  pages={1867--1887},
  year={1989}
}
%papers on Volcano plume simulation
@article{pelantinumerical,
  title={Numerical Simulation of Volcanic Jets},
  author={Pelanti, Marica and LeVeque, Randall J}
}
@article{bursik2001effect,
  title={Effect of wind on the rise height of volcanic plumes},
  author={Bursik, M},
  journal={Geophys. Res. Lett},
  volume={28},
  number={18},
  pages={3621--3624},
  year={2001}
}
@article{mastin2007user,
  title={A user-friendly one-dimensional model for wet volcanic plumes},
  author={Mastin, Larry G},
  journal={Geochemistry, Geophysics, Geosystems},
  volume={8},
  number={3},
  year={2007},
  publisher={Wiley Online Library}
}
@book{lockwood2013volcanoes,
  title={Volcanoes: global perspectives},
  author={Lockwood, John P and Hazlett, Richard W},
  year={2013},
  publisher={John Wiley \& Sons}
}
@article{wohletz1984hydrodynamic,
  title={Hydrodynamic aspects of caldera-forming eruptions: Numerical models},
  author={Wohletz, KH and McGetchin, TR and Sandford, MT and Jones, EM},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={89},
  number={B10},
  pages={8269--8285},
  year={1984},
  publisher={Wiley Online Library}
}
@article{valentine1989numerical,
  title={Numerical models of Plinian eruption columns and pyroclastic flows},
  author={Valentine, Greg A and Wohletz, Kenneth H},
  journal={J. Geophys. Res},
  volume={94},
  pages={1867--1887},
  year={1989}
}
@article{dobran1993numerical,
  title={Numerical simulation of collapsing volcanic columns},
  author={Dobran, Flavio and Neri, Augusto and Macedonio, Giovanni},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={98},
  number={B3},
  pages={4231--4259},
  year={1993},
  publisher={Wiley Online Library}
}
@article{neri1994influence,
  title={Influence of eruption parameters on the thermofluid dynamics of collapsing volcanic columns},
  author={Neri, Augusto and Dobran, Flavio},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={99},
  number={B6},
  pages={11833--11857},
  year={1994},
  publisher={Wiley Online Library}
}
@article{neri1996numerical,
  title={Numerical simulation of collapsing volcanic columns with particles of two sizes},
  author={Neri, Augusto and Macedonio, Giovanni},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={101},
  number={B4},
  pages={8153--8174},
  year={1996},
  publisher={Wiley Online Library}
}
@article{oberhuber1998volcanic,
  title={Volcanic plume simulation on large scales},
  author={Oberhuber, Josef M and Herzog, Michael and Graf, Hans-F and Schwanke, Karsten},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={29--53},
  year={1998},
  publisher={Elsevier}
}
@article{clarke2002transient,
  title={Transient dynamics of vulcanian explosions and column collapse},
  author={Clarke, AB and Voight, B and Neri, A and Macedonio, G},
  journal={Nature},
  volume={415},
  number={6874},
  pages={897--901},
  year={2002},
  publisher={Nature Publishing Group}
}
@article{neri1998role,
  title={The role of magma composition and water content in explosive eruptions: 2. Pyroclastic dispersion dynamics},
  author={Neri, Augusto and Papale, Paolo and Macedonio, Giovanni},
  journal={Journal of volcanology and geothermal research},
  volume={87},
  number={1},
  pages={95--115},
  year={1998},
  publisher={Elsevier}
}
@article{neri2002mass,
  title={Mass partition during collapsing and transitional columns by using numerical simulations},
  author={Neri, Augusto and Di Muro, Andrea and Rosi, Mauro},
  journal={Journal of volcanology and geothermal research},
  volume={115},
  number={1},
  pages={1--18},
  year={2002},
  publisher={Elsevier}
}
@article{neri2003coupled,
  title={Coupled conduit and atmospheric dispersal dynamics of the AD 79 Plinian eruption of Vesuvius},
  author={Neri, Augusto and Papale, Paolo and Del Seppia, Dario and Santacroce, Roberto},
  journal={Journal of volcanology and geothermal research},
  volume={120},
  number={1},
  pages={141--160},
  year={2003},
  publisher={Elsevier}
}
@article{neri2003multiparticle,
  title={Multiparticle simulation of collapsing volcanic columns and pyroclastic flow},
  author={Neri, Augusto and Esposti Ongaro, Tomaso and Macedonio, Giovanni and Gidaspow, Dimitri},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={108},
  number={B4},
  year={2003},
  publisher={Wiley Online Library}
}
%PADC
@article{neri20074d,
  title={4D simulation of explosive eruption dynamics at Vesuvius},
  author={Neri, Augusto and Esposti Ongaro, Tomaso and Menconi, Gianluca and De'Michieli Vitturi, Mattia and Cavazzoni, Carlo and Erbacci, Giovanni and Baxter, Peter J},
  journal={Geophysical research letters},
  volume={34},
  number={4},
  year={2007},
  publisher={Wiley Online Library}
}
%I need add most recent paper on ATHAM, PADC, and Japanese Model.
%Japanese Model
@article{suzuki2005numerical,
  title={A numerical study of turbulent mixing in eruption clouds using a three-dimensional fluid dynamics model},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro and Ogawa, Masaki and Hachisu, Izumi},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={110},
  number={B8},
  year={2005},
  publisher={Wiley Online Library}
}

@book{fischer2013mixing,
  title={Mixing in inland and coastal waters},
  author={Fischer, Hugo B and List, John E and Koh, C Robert and Imberger, Jorg and Brooks, Norman H},
  year={2013},
  publisher={Elsevier}
}

@article{suzuki2010numerical,
  title={Numerical determination of the efficiency of entrainment in volcanic eruption columns},
  author={Suzuki, YJ and Koyaguchi, T},
  journal={Geophysical Research Letters},
  volume={37},
  number={5},
  year={2010},
  publisher={Wiley Online Library}
}
@article{suzuki20133d,
  title={3D numerical simulation of volcanic eruption clouds during the 2011 Shinmoe-dake eruptions},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro},
  journal={Earth, Planets and Space},
  volume={65},
  number={6},
  pages={581--589},
  year={2013},
  publisher={Springer}
}
@article{suzuki2009three,
  title={A three-dimensional numerical simulation of spreading umbrella clouds},
  author={Suzuki, YJ and Koyaguchi, T},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={114},
  number={B3},
  year={2009},
  publisher={Wiley Online Library}
}
@article{koyaguchi2010effects,
  title={Effects of the crater on eruption column dynamics},
  author={Koyaguchi, Takehiro and Suzuki, Yujiro J and Kozono, Tomofumi},
  journal={Journal of Geophysical Research: Solid Earth (1978--2012)},
  volume={115},
  number={B7},
  year={2010},
  publisher={Wiley Online Library}
}
@article{koyaguchi2009effect,
  title={The effect of intensity of turbulence in umbrella cloud on tephra dispersion during explosive volcanic eruptions: Experimental and numerical approaches},
  author={Koyaguchi, Takehiro and Ochiai, Kiyokatsu and Suzuki, Yujiro J},
  journal={Journal of Volcanology and Geothermal Research},
  volume={186},
  number={1},
  pages={68--78},
  year={2009},
  publisher={Elsevier}
}
@article{suzuki20123,
  title={3-D numerical simulations of eruption column collapse: Effects of vent size on pressure-balanced jet/plumes},
  author={Suzuki, Yujiro J and Koyaguchi, Takehiro},
  journal={Journal of Volcanology and Geothermal Research},
  volume={221},
  pages={1--13},
  year={2012},
  publisher={Elsevier}
}

%ATHAM 
@article{oberhuber1998volcanic,
  title={Volcanic plume simulation on large scales},
  author={Oberhuber, Josef M and Herzog, Michael and Graf, Hans-F and Schwanke, Karsten},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={29--53},
  year={1998},
  publisher={Elsevier}
}
@article{herzog1998effect,
  title={The effect of phase changes of water on the development of volcanic plumes},
  author={Herzog, Michael and Graf, Hans-F and Textor, Christiane and Oberhuber, Josef M},
  journal={Journal of Volcanology and Geothermal Research},
  volume={87},
  number={1},
  pages={55--74},
  year={1998},
  publisher={Elsevier}
}
@article{langmann1997radiative,
  title={Radiative forcing of sulfate aerosols as determined by a regional circulation-chemistry transport model},
  author={Langmann, B{\"a}rbel and Herzog, Michael and Graf, Hans-F},
  journal={Journal of Aerosol Science},
  volume={28},
  pages={S389--S390},
  year={1997},
  publisher={Pergamon}
}
@article{trentmann2002simulation,
  title={Simulation of a biomass-burning plume: Comparison of model results with observations},
  author={Trentmann, J and Andreae, MO and Graf, H-F and Hobbs, PV and Ottmar, RD and Trautmann, T},
  journal={Journal of Geophysical Research: Atmospheres (1984--2012)},
  volume={107},
  number={D2},
  pages={AAC--5},
  year={2002},
  publisher={Wiley Online Library}
}
@article{textor2003injection,
  title={Injection of gases into the stratosphere by explosive volcanic eruptions},
  author={Textor, Christiane and Graf, Hans-F and Herzog, Michael and Oberhuber, JM},
  journal={Journal of Geophysical Research: Atmospheres (1984--2012)},
  volume={108},
  number={D19},
  year={2003},
  publisher={Wiley Online Library}
}
@article{herzog2003prognostic,
  title={A prognostic turbulence scheme for the nonhydrostatic plume model ATHAM},
  author={Herzog, Michael and Oberhuber, Josef M and Graf, Hans-F},
  journal={Journal of the atmospheric sciences},
  volume={60},
  number={22},
  pages={2783--2796},
  year={2003}
}

@article{textor2006volcanic1,
  title={Volcanic particle aggregation in explosive eruption columns. Part I: Parameterization of the microphysics of hydrometeors and ash},
  author={Textor, Christiane and Graf, Hans-F and Herzog, Michael and Oberhuber, Josef M and Rose, William I and Ernst, Gerald GJ},
  journal={Journal of Volcanology and Geothermal Research},
  volume={150},
  number={4},
  pages={359--377},
  year={2006},
  publisher={Elsevier}
}
@article{textor2006volcanic2,
  title={Volcanic particle aggregation in explosive eruption columns. Part II: Numerical experiments},
  author={Textor, C and Graf, Hans-F and Herzog, Michael and Oberhuber, Josef M and Rose, William I and Ernst, GGJ},
  journal={Journal of Volcanology and Geothermal Research},
  volume={150},
  number={4},
  pages={378--394},
  year={2006},
  publisher={Elsevier}
}
@book{kobs2009modeling,
  title={Modeling particle motion and near-vent deposition in explosive volcanic eruptions},
  author={Kobs, Shannon},
  year={2009},
  publisher={State University of New York at Buffalo}
}
@article{herzog2010applying,
  title={Applying the three-dimensional model ATHAM to volcanic plumes: Dynamic of large co-ignimbrite eruptions and associated injection heights for volcanic gases},
  author={Herzog, Michael and Graf, Hans-F},
  journal={Geophysical Research Letters},
  volume={37},
  number={19},
  year={2010},
  publisher={Wiley Online Library}
}
@article{bursik2009volcanic,
  title={Volcanic plumes and wind: Jetstream interaction examples and implications for air traffic},
  author={Bursik, MI and Kobs, SE and Burns, A and Braitseva, OA and Bazanova, LI and Melekestsev, IV and Kurbatov, A and Pieri, DC},
  journal={Journal of Volcanology and Geothermal Research},
  volume={186},
  number={1},
  pages={60--67},
  year={2009},
  publisher={Elsevier}
}
@article{van2015hail,
  title={Hail formation triggers rapid ash aggregation in volcanic plumes},
  author={Van Eaton, Alexa R and Mastin, Larry G and Herzog, Michael and Schwaiger, Hans F and Schneider, David J and Wallace, Kristi L and Clarke, Amanda B},
  journal={Nature communications},
  volume={6},
  year={2015},
  publisher={Nature Publishing Group}
}

%review type papers on Volcano
@article{bonadonna2012future,
  title={Future developments in modelling and monitoring of volcanic ash clouds: outcomes from the first IAVCEI-WMO workshop on Ash Dispersal Forecast and Civil Aviation},
  author={Bonadonna, Costanza and Folch, Arnau and Loughlin, Susan and Puempel, Herbert},
  journal={Bulletin of volcanology},
  volume={74},
  number={1},
  pages={1--10},
  year={2012},
  publisher={Springer}
}

%exmperimental or observation type of paper on volcano plume
@article{chojnicki2014flow,
  title={The flow structure of jets from transient sources and implications for modeling short-duration explosive volcanic eruptions},
  author={Chojnicki, KN and Clarke, AB and Adrian, RJ and Phillips, JC},
  journal={Geochemistry, Geophysics, Geosystems},
  volume={15},
  number={12},
  pages={4831--4845},
  year={2014},
  publisher={Wiley Online Library}
}

%1D plume softwares
@misc { 114,
	title = {puffin},
	month = {Jan},
	url = {https://vhub.org/resources/114},
	year = {2013},
	author = {Bursik , Marcus I and Carbonara , Alejandro Uriel and Zawicki , Sean Michael}
}
@misc { 267,
	title = {bent: A model of plumes in crossflow},
	month = {Nov},
	url = {https://vhub.org/resources/267},
	year = {2010},
	author = {Bursik , Marcus I}
}
@misc { 1194,
	title = {Plumeria 2.3.1},
	month = {Dec},
	url = {https://vhub.org/resources/1194},
	year = {2011},
	author = {Mastin , Larry G}
}
@misc { 3541,
	title = {PlumeMoM},
	month = {Mar},
	url = {https://vhub.org/resources/3541},
	year = {2015},
	author = {Vitturi , Mattia de' Michieli}
}
@misc{draxler2015hysplit,
  title={HYSPLIT (HYbrid Single-Particle Lagrangian Integrated Trajectory) Model access via NOAA ARL READY Website (). NOAA Air Resources Laboratory, Silver Spring},
  author={Draxler, R.R. and Rolph, G.D.},
  year={2015},
  publisher={MD}
}
@misc{rolph2015ready,
  title={Real-time Environmental Applications and Display sYstem (READY) Website (http://ready.arl.noaa.gov). NOAA Air Resources Laboratory, Silver Spring,},
  author={Rolph, G.D.},
  year={2015},
  publisher={MD}
}

%SPH
@article{gingold1977smoothed,
  title={Smoothed particle hydrodynamics: theory and application to non-spherical stars},
  author={Gingold, Robert A and Monaghan, Joseph J},
  journal={Monthly notices of the royal astronomical society},
  volume={181},
  number={3},
  pages={375--389},
  year={1977},
  publisher={Oxford University Press}
}
@article{lucy1977numerical,
  title={A numerical approach to the testing of the fission hypothesis},
  author={Lucy, Leon B},
  journal={The astronomical journal},
  volume={82},
  pages={1013--1024},
  year={1977}
}
@article{monaghan1992smoothed,
  title={Smoothed particle hydrodynamics},
  author={Monaghan, Joe J},
  journal={Annual review of astronomy and astrophysics},
  volume={30},
  pages={543--574},
  year={1992}
}
@article{cleary1999conduction,
  title={Conduction modelling using smoothed particle hydrodynamics},
  author={Cleary, Paul W and Monaghan, Joseph J},
  journal={Journal of Computational Physics},
  volume={148},
  number={1},
  pages={227--264},
  year={1999},
  publisher={Elsevier}
}
@article{monaghan2005smoothed,
  title={Smoothed particle hydrodynamics},
  author={Monaghan, Joe J},
  journal={Reports on progress in physics},
  volume={68},
  number={8},
  pages={1703},
  year={2005},
  publisher={IOP Publishing}
}
@article{rosswog2009astrophysical,
  title={Astrophysical smooth particle hydrodynamics},
  author={Rosswog, Stephan},
  journal={New Astronomy Reviews},
  volume={53},
  number={4},
  pages={78--104},
  year={2009},
  publisher={Elsevier}
}
@article{ellero2012smoothed,
  title={Smoothed Particle Hydrodynamics},
  author={Ellero, M},
  journal={Fluid-Structure Interactions in Soft-Matter Systems: From the Mesoscale to the Macroscale},
  pages={22},
  year={2012}
}
@article{monaghan2012smoothed,
  title={Smoothed particle hydrodynamics and its diverse applications},
  author={Monaghan, JJ},
  journal={Annual Review of Fluid Mechanics},
  volume={44},
  pages={323--346},
  year={2012},
  publisher={Annual Reviews}
}
@article{saitoh2013density,
  title={A density-independent formulation of smoothed particle hydrodynamics},
  author={Saitoh, Takayuki R and Makino, Junichiro},
  journal={The Astrophysical Journal},
  volume={768},
  number={1},
  pages={44},
  year={2013},
  publisher={IOP Publishing}
}
@techreport{takayuki2012density,
  title={A Density Independent Formulation of Smoothed Particle Hydrodynamics},
  author={Takayuki, Saitoh R and Makino, Junichiro},
  year={2012}
}
@article{price2012smoothed,
  title={Smoothed particle hydrodynamics and magnetohydrodynamics},
  author={Price, Daniel J},
  journal={Journal of Computational Physics},
  volume={231},
  number={3},
  pages={759--794},
  year={2012},
  publisher={Elsevier}
}
@article{monaghan1985extrapolating,
  title={Extrapolating B splines for interpolation},
  author={Monaghan, JJ},
  journal={Journal of Computational Physics},
  volume={60},
  number={2},
  pages={253--262},
  year={1985},
  publisher={Elsevier}
}
@article{monaghan1995sph,
  title={SPH simulation of multi-phase flow},
  author={Monaghan, JJ and Kocharyan, A},
  journal={Computer Physics Communications},
  volume={87},
  number={1},
  pages={225--235},
  year={1995},
  publisher={Elsevier}
}
@article{morris1997modeling,
  title={Modeling low Reynolds number incompressible flows using SPH},
  author={Morris, Joseph P and Fox, Patrick J and Zhu, Yi},
  journal={Journal of computational physics},
  volume={136},
  number={1},
  pages={214--226},
  year={1997},
  publisher={Elsevier}
}

@article{chen1999improvement,
  title={An improvement for tensile instability in smoothed particle hydrodynamics},
  author={Chen, JK and Beraun, JE and Jih, CJ},
  journal={Computational Mechanics},
  volume={23},
  number={4},
  pages={279--287},
  year={1999},
  publisher={Springer}
}
@article{adami2012generalized,
  title={A generalized wall boundary condition for smoothed particle hydrodynamics},
  author={Adami, S and Hu, XY and Adams, NA},
  journal={Journal of Computational Physics},
  volume={231},
  number={21},
  pages={7057--7075},
  year={2012},
  publisher={Elsevier}
}
@article{kumar2013parallel,
  title={Parallel Godunov smoothed particle hydrodynamics (SPH) with improved treatment of Boundary Conditions and an application to granular flows},
  author={Kumar, Dinesh and Patra, Abani K and Pitman, E Bruce and Chi, H},
  journal={Computer Physics Communications},
  volume={184},
  number={10},
  pages={2277--2286},
  year={2013},
  publisher={Elsevier}
}
@article{monaghan2009sph,
  title={SPH particle boundary forces for arbitrary boundaries},
  author={Monaghan, Joe J and Kajtar, Jules B},
  journal={Computer Physics Communications},
  volume={180},
  number={10},
  pages={1811--1820},
  year={2009},
  publisher={Elsevier}
}
@article{yildiz2009sph,
  title={SPH with the multiple boundary tangent method},
  author={Yildiz, M and Rook, RA and Suleman, A},
  journal={International journal for numerical methods in engineering},
  volume={77},
  number={10},
  pages={1416--1438},
  year={2009},
  publisher={Wiley Online Library}
}
@inproceedings{ihmsen2011parallel,
  title={A Parallel SPH Implementation on Multi-Core CPUs},
  author={Ihmsen, Markus and Akinci, Nadir and Becker, Markus and Teschner, Matthias},
  booktitle={Computer Graphics Forum},
  volume={30},
  number={1},
  pages={99--112},
  year={2011},
  organization={Wiley Online Library}
}
@article{herault2010sph,
  title={Sph on gpu with cuda},
  author={H{\'e}rault, Alexis and Bilotta, Giuseppe and Dalrymple, Robert A},
  journal={Journal of Hydraulic Research},
  volume={48},
  number={S1},
  pages={74--79},
  year={2010},
  publisher={Taylor \& Francis}
}
@article{crespo2011gpus,
  title={GPUs, a new tool of acceleration in CFD: efficiency and reliability on smoothed particle hydrodynamics methods},
  author={Crespo, AJC and Dominguez, Jose M and Barreiro, Anxo and G{\'o}mez-Gesteira, Moncho and Rogers, Benedict D},
  journal={PLoS One},
  volume={6},
  number={6},
  pages={e20685},
  year={2011},
  publisher={Public Library of Science}
}
@article{dominguez2013optimization,
  title={Optimization strategies for CPU and GPU implementations of a smoothed particle hydrodynamics method},
  author={Dom{\'\i}nguez, Jose M and Crespo, Alejandro JC and G{\'o}mez-Gesteira, Moncho},
  journal={Computer Physics Communications},
  volume={184},
  number={3},
  pages={617--627},
  year={2013},
  publisher={Elsevier}
}
@article{loren2014two,
  title={Two-fluid dust and gas mixtures in smoothed particle hydrodynamics: a semi-implicit approach},
  author={Lor{\'e}n-Aguilar, Pablo and Bate, Matthew R},
  journal={Monthly Notices of the Royal Astronomical Society},
  volume={443},
  number={1},
  pages={927--945},
  year={2014},
  publisher={Oxford University Press}
}
@article{rook2007modeling,
  title={Modeling transient heat transfer using SPH and implicit time integration},
  author={Rook, Rusty and Yildiz, Mehmet and Dost, Sadik},
  journal={Numerical Heat Transfer, Part B: Fundamentals},
  volume={51},
  number={1},
  pages={1--23},
  year={2007},
  publisher={Taylor \& Francis}
}

@article{vacondio2012accurate,
  title={Accurate particle splitting for smoothed particle hydrodynamics in shallow water with shock capturing},
  author={Vacondio, R and Rogers, BD and Stansby, PK},
  journal={International Journal for Numerical Methods in Fluids},
  volume={69},
  number={8},
  pages={1377--1410},
  year={2012},
  publisher={Wiley Online Library}
}
@article{feldman2007dynamic,
  title={Dynamic refinement and boundary contact forces in SPH with applications in fluid flow problems},
  author={Feldman, J and Bonet, J},
  journal={International Journal for Numerical Methods in Engineering},
  volume={72},
  number={3},
  pages={295--324},
  year={2007},
  publisher={Wiley Online Library}
}
@article{lastiwka2009permeable,
  title={Permeable and non-reflecting boundary conditions in SPH},
  author={Lastiwka, Martin and Basa, Mihai and Quinlan, Nathan J},
  year={2009},
  publisher={Wiley}
}
@article{colagrossi2005meshless,
  title={A meshless Lagrangian method for free-surface and interface flows with fragmentation},
  author={Colagrossi, Andrea},
  year={2005},
  publisher={La Sapienza}
}
@article{molteni2009simple,
  title={A simple procedure to improve the pressure evaluation in hydrodynamic context using the SPH},
  author={Molteni, Diego and Colagrossi, Andrea},
  journal={Computer Physics Communications},
  volume={180},
  number={6},
  pages={861--872},
  year={2009},
  publisher={Elsevier}
}
@article{kondo2011improvement,
  title={Improvement of stability in moving particle semi-implicit method},
  author={Kondo, Masahiro and Koshizuka, Seiichi},
  journal={International Journal for Numerical Methods in Fluids},
  volume={65},
  number={6},
  pages={638--654},
  year={2011},
  publisher={Wiley Online Library}
}
@article{macia2011theoretical,
  title={Theoretical analysis of the no-slip boundary condition enforcement in SPH methods},
  author={Maci{\'a}, Fabricio and Antuono, Matteo and Gonz{\'a}lez, Leo M and Colagrossi, Andrea},
  journal={Progress of theoretical physics},
  volume={125},
  number={6},
  pages={1091--1121},
  year={2011},
  publisher={Oxford University Press}
}
@article{gotoh2004sph,
  title={SPH-LES model for numerical investigation of wave interaction with partially immersed breakwater},
  author={Gotoh, Hitoshi and Shao, Songdong and Memita, Tetsu},
  journal={Coastal Engineering Journal},
  volume={46},
  number={01},
  pages={39--63},
  year={2004},
  publisher={World Scientific}
}
@article{shao2006sph,
  title={SPH computation of plunging waves using a 2-D sub-particle scale (SPS) turbulence model},
  author={Shao, Songdong and Ji, Changming},
  journal={International Journal for numerical methods in fluids},
  volume={51},
  number={8},
  pages={913--936},
  year={2006},
  publisher={Wiley Online Library}
}
@article{hu2009constant,
  title={A constant-density approach for incompressible multi-phase SPH},
  author={Hu, XY and Adams, NA},
  journal={Journal of Computational Physics},
  volume={228},
  number={6},
  pages={2082--2091},
  year={2009},
  publisher={Elsevier}
}

@article{tong2014incompressible,
  title={An incompressible multi-phase smoothed particle hydrodynamics (SPH) method for modelling thermocapillary flow},
  author={Tong, Mingming and Browne, David J},
  journal={International Journal of Heat and Mass Transfer},
  volume={73},
  pages={284--292},
  year={2014},
  publisher={Elsevier}
}
%GSPH
@article{inutsuka2002reformulation,
  title={Reformulation of smoothed particle hydrodynamics with Riemann solver},
  author={Inutsuka, Shu-ichiro},
  journal={Journal of Computational Physics},
  volume={179},
  number={1},
  pages={238--267},
  year={2002},
  publisher={Elsevier}
}
@inproceedings{borgani2012hydrodynamic,
  title={Hydrodynamic Simulations with the Godunov SPH},
  author={Borgani, S and Murante, G and Brunino, R and Cha, S-H},
  booktitle={Advances in Computational Astrophysics: Methods, Tools, and Outcome},
  volume={453},
  pages={259},
  year={2012}
}

%%Reformulate SPH by Galerkin
@article{diltsg1999moving,
  title={Moving-least-squares-particle hydrodynamics—I Consistency and stability},
  author={Dilts, Gary A},
  journal={International Journal for Numerical Methods in Engineering},
  volume={44},
  number={8},
  pages={1115--1155},
  year={1999},
  publisher={Wiley Online Library}
}

@article{monaghan1985refined,
  title={A refined particle method for astrophysical problems},
  author={Monaghan, Joseph J and Lattanzio, John C},
  journal={Astronomy and astrophysics},
  volume={149},
  pages={135--143},
  year={1985}
}

%%turbulent && SPH
@article{shao2006sph,
  title={SPH computation of plunging waves using a 2-D sub-particle scale (SPS) turbulence model},
  author={Shao, Songdong and Ji, Changming},
  journal={International Journal for numerical methods in fluids},
  volume={51},
  number={8},
  pages={913--936},
  year={2006},
  publisher={Wiley Online Library}
}
@article{violeau2007numerical,
  title={Numerical modelling of complex turbulent free-surface flows with the SPH method: an overview},
  author={Violeau, Damien and Issa, Reza},
  journal={International Journal for Numerical Methods in Fluids},
  volume={53},
  pages={277--304},
  year={2007}
}

@article{holm1999fluctuation,
  title={Fluctuation effects on 3D Lagrangian mean and Eulerian mean fluid motion},
  author={Holm, Darryl D},
  journal={Physica D: Nonlinear Phenomena},
  volume={133},
  number={1},
  pages={215--269},
  year={1999},
  publisher={Elsevier}
}

@article{monaghan2011turbulence,
  title={A turbulence model for Smoothed Particle Hydrodynamics},
  author={Monaghan, JJ},
  journal={European Journal of Mechanics-B/Fluids},
  volume={30},
  number={4},
  pages={360--370},
  year={2011},
  publisher={Elsevier}
}

@article{violeau2007numerical,
  title={Numerical modelling of complex turbulent free-surface flows with the SPH method: an overview},
  author={Violeau, Damien and Issa, Reza},
  journal={International Journal for Numerical Methods in Fluids},
  volume={53},
  pages={277--304},
  year={2007}
}

@article{issa2005numerical,
  title={Numerical assessment of the Smoothed Particle Hydrodynamics gridless method for incompressible flows and its extension to turbulent flows},
  author={Issa, Reza},
  year={2005}
}

@article{monaghan2002sph,
  title={SPH compressible turbulence},
  author={Monaghan, JJ},
  journal={Monthly Notices of the Royal Astronomical Society},
  volume={335},
  number={3},
  pages={843--852},
  year={2002},
  publisher={Oxford University Press}
}

%@book{gatski2013compressibility,
%  title={Compressibility, turbulence and high speed flow},
%  author={Gatski, Thomas B and Bonnet, Jean-Paul},
%  year={2013},
%  publisher={Academic Press}
%}

@misc {NASACompressibleTurbulence,
	title = {Implementing Turbulence Models into the Compressible RANS Equations},
	month = {Agu},
	url={http://turbmodels.larc.nasa.gov/implementrans.html},
	year = {2014},
	author = {Christopher Rumsey}
}

@book{cebeci2013analysis,
  title={Analysis of turbulent flows with computer programs},
  author={Cebeci, Tuncer},
  year={2013},
  publisher={Butterworth-Heinemann}
}

@article{kays1994turbulent,
  title={Turbulent Prandtl number—where are we?},
  author={Kays, William M},
  journal={Journal of Heat Transfer},
  volume={116},
  number={2},
  pages={284--295},
  year={1994},
  publisher={American Society of Mechanical Engineers}
}

%%SPH parallel
@article{goozee2003distributed,
  title={Distributed and shared memory parallelism with a smoothed particle hydrodynamics code},
  author={Gooz{\'e}e, Richard J and Jacobs, Peter A},
  journal={ANZIAM Journal},
  volume={44},
  pages={202--228},
  year={2003}
}

@article{holmes2011framework,
  title={A framework for parallel computational physics algorithms on multi-core: SPH in parallel},
  author={Holmes, David W and Williams, John R and Tilke, Peter},
  journal={Advances in Engineering Software},
  volume={42},
  number={11},
  pages={999--1008},
  year={2011},
  publisher={Elsevier}
}

@inproceedings{wenbo2014performance,
  title={Performance analysis of parallel smoothed particle hydrodynamics on multi-core CPUs},
  author={Wenbo, Chen and Yao, Yucheng and Zhang, Yang},
  booktitle={Cloud Computing and Internet of Things (CCIOT), 2014 International Conference on},
  pages={85--90},
  year={2014},
  organization={IEEE}
}

@inproceedings{adams2007adaptively,
  title={Adaptively sampled particle fluids},
  author={Adams, Bart and Pauly, Mark and Keiser, Richard and Guibas, Leonidas J},
  booktitle={ACM Transactions on Graphics (TOG)},
  volume={26},
  number={3},
  pages={48},
  year={2007},
  organization={ACM}
}

@article{dominguez2011optimization,
  title={Optimization strategies for parallel CPU and GPU implementations of a meshfree particle method},
  author={Dom{\'\i}nguez, Jose M and Crespo, Alejandro JC and G{\'o}mez-Gesteira, Moncho},
  journal={arXiv preprint arXiv:1110.3711},
  year={2011}
}

@article{ferrari2009new,
  title={A new 3D parallel SPH scheme for free surface flows},
  author={Ferrari, Angela and Dumbser, Michael and Toro, Eleuterio F and Armanini, Aronne},
  journal={Computers \& Fluids},
  volume={38},
  number={6},
  pages={1203--1217},
  year={2009},
  publisher={Elsevier}
}

@article{crespo2015dualsphysics,
  title={DualSPHysics: Open-source parallel CFD solver based on Smoothed Particle Hydrodynamics (SPH)},
  author={Crespo, AJC and Dom{\'\i}nguez, JM and Rogers, BD and G{\'o}mez-Gesteira, M and Longshaw, S and Canelas, R and Vacondio, R and Barreiro, A and Garc{\'\i}a-Feal, O},
  journal={Computer Physics Communications},
  volume={187},
  pages={204--216},
  year={2015},
  publisher={Elsevier}
}

%Dr. Patra's paper
@article{laszloffy2000simple,
  title={Simple data management, scheduling and solution strategies for managing the irregularities in parallel adaptive hp finite element simulations},
  author={Laszloffy, Andras and Long, Jingping and Patra, Abani K},
  journal={Parallel Computing},
  volume={26},
  number={13},
  pages={1765--1788},
  year={2000},
  publisher={Elsevier}
}

@article{patra2003data,
  title={Data structures and load balancing for parallel adaptive hp finite-element methods},
  author={Patra, AK and Laszloffy, A and Long, J},
  journal={Computers \& Mathematics with Applications},
  volume={46},
  number={1},
  pages={105--123},
  year={2003},
  publisher={Elsevier}
}

@article{patra2005parallel,
  title={Parallel adaptive numerical simulation of dry avalanches over natural terrain},
  author={Patra, Abani K and Bauer, AC and Nichita, CC and Pitman, E Bruce and Sheridan, MF and Bursik, M and Rupp, B and Webber, A and Stinton, AJ and Namikawa, LM and others},
  journal={Journal of Volcanology and Geothermal Research},
  volume={139},
  number={1},
  pages={1--21},
  year={2005},
  publisher={Elsevier}
}

@article{pitman2003computing,
  title={Computing granular avalanches and landslides},
  author={Pitman, E Bruce and Nichita, C Camil and Patra, Abani and Bauer, Andy and Sheridan, Michael and Bursik, Marcus},
  journal={Physics of Fluids (1994-present)},
  volume={15},
  number={12},
  pages={3638--3646},
  year={2003},
  publisher={AIP Publishing}
}

@article{patra1995problem,
  title={Problem decomposition for adaptive hp finite element methods},
  author={Patra, Abani and Oden, J Tinsley},
  journal={Computing Systems in Engineering},
  volume={6},
  number={2},
  pages={97--109},
  year={1995},
  publisher={Elsevier}
}

@inproceedings{patra1999efficient,
  title={Efficient mesh partitioning for adaptive hp finite element meshes},
  author={Patra, AK and Kim, DW},
  booktitle={In International Conference on Domain Decomposition Methods},
  year={1999},
  organization={Citeseer}
}

%SFC 
@book{sagan2012space,
  title={Space-filling curves},
  author={Sagan, Hans},
  year={2012},
  publisher={Springer Science \& Business Media}
}

%domain partition
@incollection{biswas1999experiments,
  title={Experiments with repartitioning and load balancing adaptive meshes},
  author={Biswas, Rupak and Oliker, Leonid},
  booktitle={Grid Generation and Adaptive Algorithms},
  pages={89--111},
  year={1999},
  publisher={Springer}
}

### 2015-12-21 Paper for submission.tex
updated the performance results of the most recent, paper format is in required style

%% bare_conf_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE Computer
%% Society conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference,compsoc]{IEEEtran}
% Some/most Computer Society conferences require the compsoc mode option,
% but others may want the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
%\documentclass[conference,compsoc{/Users/czxpc/Documents/shared/report/Report/latex/Sty/IEEEtran}

%Added by Zhixuan Cao to make reference more compact
\def\IEEEbibitemsep{0pt plus .6pt}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
 \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Data Management and Domain Adjusting Strategies for Implementing Parallel SPH Method in JPUE Simulation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra}
\IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\
 The University at Buffalo\\
 Buffalo, New York 14260\\
Email: zhixuanc@buffalo.edu, abani@buffalo.edu}
%\and
%\IEEEauthorblockN{Homer Simpson}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: homer@thesimpsons.com}
%\and
%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}
}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page (and note that there is less available width in this regard for
% compsoc conferences compared to traditional conferences), use this
% alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

%The following package is added by Zhixuan Cao, I am not sure whether should I have %double line for IEEE computer society conference or not, need to make sure
%\usepackage{hhline} %to add double horizontal line in tabular

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and decompose (redecompose) the domain. Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are managed by hashtables which ensures quick and flexible accessing, adding and deleting of element. Additional link list attached to the hashtable is used to handle hash confliction. A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance. Furtherly, computational domain is adjusted during simulation to reduce computational cost. 
%Numerical tests show that our code has linear speed up.
These strategies can be further applied to many other implementations of meshfree methods, especially those implemetations that require flexible accessing, adding and deleting of particles and backgroud grids.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
SPH is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
Subsequently it was extended to large strain solid mechanics and computational fluid mechanics. 
A lot of research has been done in parallellization of SPH. 
Goozee\cite{goozee2003distributed} implemented a simple SPH code using MPI, OpenMP and BSP. 
Chen Wenbo\cite{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
David W. Holmes\cite{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
Domínguez\cite{dominguez2011optimization} presented optimizations for both CPU and GPU parallelization of a SPH method. 
Angela Ferrari\cite{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme. 
D. Kumar and Abani\cite{kumar2013parallel} implemented CPU parallel Godunov SPH in simulation of granular flows.
A.J.C. Crespo\cite{crespo2015dualsphysics} used the parallel power computing of GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
However, most implementation of parallel SPH method is limitted to bechmark problems like dam break, or relative simple scenarios like breaking-waves, floodings ect. Rare work has been done in simulation of more complicated problems, such as eruption of volcano plume which is essentially a mutilple phases, turbulent, ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction ect. Prediction of such more complicated phenomena with acceptable accuracy at given time window can not been acomplished without parallel computing. What's more, imposing of some types of boundary conditions (such as realistic wind feild, eruption boundary condition) requires dynamiclly adding and removing of particles during simulation. This requies efficient and more flexible data management scheme. What need to point out is that the benefit of flexible data management strategies are not only limited to specific implementations of SPH. Actually, flexibility in data management are more desperately needed in some advanced techniques of SPH, such as dynamic particle splitting techniques\cite{vacondio2012accurate, feldman2007dynamic}, which will give greater resolution at the area of interest by spliiting one large particle to several smaller ones.
In this paper, we implement SPH to simulate a simplified version of complicated volcanic plume: the JPUE, and develop data management schemes for it.\\
Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing. (eg.\cite{ferrari2009new, crespo2015dualsphysics}). Less attention has been paid to developing of more flexible data management schemes for more complicated problems. 
Fortunately, efficient and flexible data management strategies for high performance computing have been successfully implemented in mesh based methods(eg. \cite{laszloffy2000simple} for adaptive hp FEM, and \cite{pitman2003computing, patra2005parallel} for FVM). Motivated by techniques developed for mesh based methods, we present a complete framwork for parallelizing SPH program with MPI standard model allowing more flexible and efficient data access in this paper.\\
Any implementation of SPH code requires efficient searching and updating of neighbors during simulation. We adopted background grid which was proposed by Monaghan and Lattanzio\cite{monaghan1985refined} and implemented widespreadly in parallel SPH. The background grid is also used for domain decomposition in SPH. We refer to the elements of background grid, namely a squares for two dimension and cubics for three dimension, as buckets. As for the storage of physical quantities associated to each particle, different strategies was adopted in existing implementations of SPH. 
In both SPHysics\cite{dominguez2011optimization} and DualSPHysics\cite{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells. This has two advantages: 1)access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known. But adding, deleting and especially accessing of particles are cumbersome. Angela Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation. Storage problems caused by fix-size arrays are thereby also eliminated. We define C++ classes which contains all data of particle and bucket. As for the storage of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list. Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier. The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have a different identifier. As for domain decomposion, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming. So we adopt an easy-programming scheme based on SFC \cite{patra1999efficient}.\\
To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluid. This feature is accomplished by adding a scan function to monitor the most outside layer of the domain and turn ghost particles to real particles at proper time. 
%The computational cost of this additional scan function is low.
\\
The data structure, particle and bucket indexing strategies, domain decomposition , dynamic load balancing method and domain adjusting strategies in this paper can be easily adopted by other implementations of any meshfree methods(include SPH). The flexibility of data accessing enables implementing of meshless methods for solving of more complicated problems and using of more advanced techniques.\\
\section{Data Structure and Load Balance}
\begin{figure}[!t]
\centering
\includegraphics[scale=0.34]{Work_flow}
\caption{Basic Work flow for SPH}
\label{fig:Work_flow}
\end{figure}
SPH is a meshfree, Lagrangian method. The domain is discretized by particles and the position of particle is updated at every time step. The physical laws (such as conservation laws of mass, momentum and energy) written in the form of partial differential equations need to be transformed into the Lagrangian formalism. Using a kernel function that provides the weighted estimate of the field variables at a discrete point (particle), the integral equations are evaluated as sums over neighbor particles. Only particles located within support of kernel function will interact. Thus, physical proterties (position, density, velocity, internal energy, pressure) are updated based on its neighbors. So a neighbor searching need to be carried out before updating of physical proterties. We use buckets which overlap the whole domain and keep fixed in time during the entire simulation, to reduce searching cost. Domain decomposition will be based on SFC going through centroids of all buckets. A basic works flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
Particles need to be added and(or) removed during simulation of JPUE problems. Wall boundary conditions in SPH can be imposed either by adding a force term or using ghost particles. We adopt the latter one in our simulation. As our computational domain will be adjusted during simulation, wall ghost particles need to be added during simulation. New particles also need to be added for the eruption boundary condition. 
We will describe in this section these strategies which allows flexible deleting and adding.
\subsection{SFC based indexing}
Our data structure starts from assigning each particle and bucket an identifier, we refer to it as key, which should be unique throughout simulation. The key for a bucket is determined by centroid coordinates of the bucket while the key for a particle is determined by adding coordinates and adding time step of the particle. The map from coordinates to key is based on SFC.\\
The SFC \cite{sagan2012space} maps n-dimentional space to a one dimentional sequence. The standard procedure for obtainning SFC is: 
\begin{itemize}
\item Scale coordinates into $[0,1]^n $ based on maximum and minimum coordinates of the computational domain: $\textbf{X}^\prime \rightarrow \textbf{X}$
\item Compute $k_r = h_n(\textbf{X})$. Where $h_n$ is the map $h_n: [0,1]^n \rightarrow [0,1]$. 
\item Convert $k_r$ to integer $k$ by multiplying $k_r$ with a very large number and removing decimal part.
\item All keys are sorted to form a sequence which is SFC. The SFC represents a curve passing through all particles (or centroid of buckets).
\end{itemize}
%In the following sections, the physics and SPH implementation of inject flow problem is given, then 
\begin{figure*}[!t]
\centering
\subfloat[SFCs passing all particles and buckets]{\includegraphics[scale=0.20]{SFC_particles_buckets}}
\label{fig:SFC_particles_buckets}
\hfil
\subfloat[Domain decomposition based on SFC of buckets]{\includegraphics[scale = 0.22]{SFC_particles_buckets_partition}}
\label{fig:SFC_particles_buckets_partition}
\caption{Spcace filling curve orderings of buckets and particles within the buckets}
\label{fig:SFC_domain_decomposition}
\end{figure*}
Scheme for constructing the map $h_n$ can be found in \cite{patra1995problem}. These keys denote a simple addressing/ordering scheme for the data and computations, i.e., a simple global index space for all the objects.\\
SFC-based indexing scheme can guarantee uniqueness of particle identifier only in simple scenarios when particles are added once while setting up initial condition. In some situations, new particles need to be added while simulation. For example, new particles need to be added at the bottom of the eject vent for JPUE simulation. To distinguish particles added at the "same place" (the small area, all points whithin which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key. The time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step. The map $h_n$ will become:
\begin{equation}
h_n: [0,1]^n \times \textbf{T} \rightarrow [0,1] \times \textbf{N}
\end{equation}
Where $\textbf{T} \subset [0,\infty)$ is the time step dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.
To guarantee locality, sorting of particle keys is majorly based on $k$, that is to say, particle with smaller $k$ always comes before particles with larger $k$. For these particles have the same $k$, ordering of them will depend on $t$. Figure \ref{fig:SFC_particles_buckets} shows SFC ordering of buckets and particles in buckets. 
Several features of such indexing scheme are suitable for SPH:
\begin{itemize}
\item Guarantee uniqueness of keys.
\item Key of each object is generated purely based on its own coordinates. When add new objects on different processes, key of each object can be generated fast and independently.
\item Objects that locate closely in the Euclidian space will also be close to each other in the one dimensional SFC key space in the mean sense. Since SPH particles only interact with its neighbors, geometric locality can be exploited for efficient storage and retrieval of bucket and particle data.
\item This type of key effectivelly generate a global address space. Globality of key and conservation of locality make it easy to partition the sorted key squence and obtain the decompositiion of the problem.
\end{itemize}
What need emphasize here is that motion of particles might mess up locality that established based on initial coordinates of particles. As particles are moving regularly, the locality of most of particles should still be conserved during simulation.
%If we can design a re-numberring method which is computationally cheap, we can restore the locality.
\subsection{Data structure}
\subsubsection{Particle and bucket}
The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbor searching and domain decomposition. Both are defined as classes in C++. Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, like density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg. pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is vector of particle keys in our application). Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (27 neighbor buckets including itself for three dimension and 9 for two dimension) and possessed particles. Objects defined based on these two classes are then accesse through hash tables.
\subsubsection{Hash table and hash confliction}
As discussed at the beginning of this section, implementation of SPH in more realistic scenarios requires dynamic memory management and flexible data access. One of the fundemental data structures that satisfy such requirement is hash table. Another option is B-tree. We adopt hash table. An implementation of B-tree under a similar situation for mesh based methods can be found in other papers(eg.\cite{patra2003data}).\\
Hash table, which is divided into slots, are array based data structure. Based on the key, the address-calculator(hashing) function determines in which slot the data should be stored. The hashing function maps from key to the slot index:
\begin{equation}
slot\,index = hash(key)
\end{equation}
The hash table has O(1) data accessing, adding and deleting properties when there is no confliction. How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table. As the distribution of keys is determined by particles' initial location (or centroids of buckets), the hash table size is under our control. We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency. Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions. Abani\cite{patra2003data} did numerical experiments to make trade-off to determine hash table size.\\
 %We should also do some experiment to optimize our hash table size.
One way to handle hash confliction is using an additional sorted vector attached to the hash table. When several keys hash to the same slot, a vector will be created. The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrievaling . Another option to handle hash confliction is using an additional link list which is more flexible in memory allocation. The average time complexity of binary search is O(log n) while that for linear search based on link list is O(n). However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longger. Choosing of proper way to handle hash confliction greatly depends on the problem itself. For the test problem in this paper, successively adding of particles at the bottom of the eject vent will lead to hash confliction of many particles, which implies a long link list. But considering the very long conflictions only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash confliction. This decision was made based on numerical experiments.\\
\subsubsection{Hash function}
\begin{figure}[!t]
\centering
\includegraphics[scale=0.255]{Particle_adding_with_link}
\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, these new particles will be stored in external link list}
\label{fig:Particle_adding_with_link}
\end{figure} 
For time-independent keys, the hash function can be a simple function like:
\begin{equation}
\label{eq:hash_function}
\begin{split}
Slot\,Index= &\frac{Key - Min\,Key}{Max\,Key - Min\,Key} \\
& \times Hash\,Table\,Size 
\end{split}
\end{equation}
One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as higher digit and $t$ as lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent. That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}. To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}). 
\subsection{Load balance strategy}
\subsubsection{Weighted work load}
Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}). As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data. Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now becomes a weighted sequence. Domain decomposition will conducted based on the weighted SFC of buckets.
\begin{table}[t!]
  \renewcommand{\arraystretch}{1.2}
  \centering
  \caption{Computational Cost for Different Steps Per Particle}
  \label{tab:Computational_cost_steps}
  \begin{tabular}{|c|c|c|}
    \hline
    Step & Cost ($ms$) & Abbreviation\\
    %\hhline{|=|=|=|}
    \hline
    neighbor search & 0.41 & NS\\
    \hline
    update momentum and energy & 0.70 & UPME\\
    \hline
    update density & 0.42 & UPD\\
    \hline
    update position & 0.02 & UPP\\
    \hline
    velocity filtering& 0.43 & VF\\
    \hline
    apply wall bc & 0.75 & WBC\\
    \hline
  \end{tabular}
\end{table}
\begin{table}[t!]
  \renewcommand{\arraystretch}{1.2}
  \centering
  \caption{Computational Work Load for Each Type of Particle}
  \label{tab:Computational_cost_particles}
  \begin{tabular}{|c|r|r|r|r|r|r|r|}
    \hline
    Particle type & NS & UPME & UPD & UPP &VF &WBC &Total\\
    \hline
    %\hhline{|=|=|=|=|=|=|=|=|}
    Real & Yes & Yes & Yes & Yes & Yes & No & 2.00\\
    \hline
    wall ghost & No & No & No & No & No & Yes &0.75\\
    \hline
    eruption ghost & No & No & No & Yes & No & No & 0.02\\
    \hline
    pressure ghost & No & No & No & No & No & No & 0.00\\
    \hline
  \end{tabular}
\end{table}
\subsubsection{Domain decomposition and dynamic load balancing}
Domain decomposition will be conducted based the weighted SFC of buckets. Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets. The particles are automatically split into several groups along with buckets that contain them. As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition. A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}. A scalability comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger. 
%Some parallel implementations of SPH use 2D decomposition \cite{kumar2013parallel} even 1D decomposition \cite{marrone2012study, cherfils2012josephine}.  
%In our current implementation, the domain is partitioned only in $x-y$ plane. A two dimensional footprint of the three dimensional domain(as well as buckets) is generated by projecting. 
%SFC for the two dimensional footprint is partitioned for domain decomposition. The work load of each object on the SFC of footprint buckets in $x-y$ plane will be the summation of work loads of all buckets above the footprint. 
Movement of particles, adding of new particles, adjusting of domain will lead to important load imbalance between processes. To handle this, computational load is monitored at given interval (The interval is optimized based on numerical experiments). Repartitioning is carried out if load imbalance is larger than a given tolenrance. \\
%Figure \ref{fig:t0_domain_decomposition} and \ref{fig:t10_domain_decomposition} show the initial decomposition and domain decomposion at 10 seconds.\\

%We adopt the same SFC based partition strategy as described by Patra \ref{patra1999efficient}.

%\begin{figure}[h]
%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: initial domain decomposition}
%\centering
%\label{fig:t0_domain_decomposition}
%\includegraphics[width=0.5\textwidth]{t0_domain_decomposition}
%\end{figure}
%\begin{figure}[h]
%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: decomposition at t=10s}
%\centering
%\label{fig:t10_domain_decomposition}
%\includegraphics[width=0.5\textwidth]{t10_domain_decomposition}
%\end{figure}
As some of the neighbor particles reside in other
partitions. A set of “guest” particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 
\section{Adjusting of Domain During Simulation}
%SPH particles can be viewed as discretization points\cite{diltsg1999moving}. 
As a Lagrangian method, SPH is able to automatically adjust computational domain as the postition of the dicretization points are updated at every time step. However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone. Because the whole domain, which occupied by stationary fluid before eject fluids reaching there, has to be discretized at the very beginning of simulation. A lot of CPU time will be spent on computing of "stationary" particles. It is pure wasting of computational resources. If simulating of stationary particls can be avoided, the computational cost will be reduced greatly. To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. We propose a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\
\begin{figure}[h]
\caption{Work flow that enables domain adjusting feature}
\centering
\label{fig:Work_flow_adjust}
\includegraphics[width=0.5\textwidth]{Work_flow_adjust}
\end{figure}
The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class. Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0). The involved particles are particles that have already been involved in eject mixing. The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles. So they will be involved in the mixing process in the near future. These not involved particles are particles that far away from the eject source and are still at the initial state. For these not involved particles, it is meaningless to update its physical properties. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary. That is to say, only potential involved and involved particles need to be simulated.
As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced. When originally stationary air is influenced by erupted material, the mass fraction of the erupted material of mixture will increase from zero to a positive value. So we can determine whether a particle is involved or not is based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation). Other physical properties, such velocity, can also serve as alternative "swich criteria".\\
\begin{figure*}[!t]
\centering
\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.24]{Involved_initial}}
\label{fig:Involved_initial}
\hfil
\subfloat[Computational domain at $t=t0 + \delta t$]{\includegraphics[scale = 0.24]{Involved_domain_adjusting}}
\label{fig:Involved_domain_adjusting}
\caption{Domain adjusting based on Involved flag of particles}
\label{fig:Domain_adjusting}
\end{figure*}
A has-involved-flag is added to bucket class, too. Buckets are also categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And its neighbor buckets that do not contain any involved particle will be set to be has potential involved. 
All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles. There are two situations that will switch a has potential involved bucket to be a has involved bucket. First, when a involved particle enters a has potential involved bucket. Second, when mass fraction of a particle (It should be a potental involved particle) exceeds the threshold, this particle will turn to a involved particles and if this is the first involved particle in the bucket, the bucket will turn from a has potential involved bucket to a has involved bucekt.\\
The most outside buckets layer of all has potential involved buckets will be scaned every time after updating. If any of these buckets becomes has involved, the domain will be enlarged by turnning the pressure ghost particles to real particles and switching involved-flag from 0 to 1. Also, the buckets that originally contain pressure ghost particles will become has potential involved. New pressure ghost particles and wall ghost particles will be added around the adjusted domain.
This domain adjusting process is shown in figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}.
The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}\\
For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption. When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$. This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation. Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
\section{Numerical Test}
As we are targeting at developing data management and paralell strategies for more complicated implementations of SPH which demmand quick and flexible data access, delete and add. The test problem should have such demand. Volcanic eruption which is essentially a mutilple phases, turbulent, ejection mixing flow accompanied with microphysics processes requires more flexibile data management. We adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem. In this model, one phase is air while another phase is ejected material.
\subsection{Governing equations\cite{suzuki2005numerical} and boundary conditions}
Based on Navier-Stokes equations and several simplification assumption, the governing equations in Eulerian form are:
%the following assumptions are made:
%\begin{itemize}
%\item Molecular Viscosity is neglected since eddy viscosity due to turbulence is dominant. As heat conduction is much smaller than turbulent heat exchange \cite{oberhuber1998volcanic}, We only consider turbulent heat exchange.
%\item Erupted material consist of solid with different size and mixture of gas (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2). Air (also a mixture of different gas) is assumed to be another phase (phase 1).
%\item We assume immediate thermodynamic equilibrium so that no separate energy equation is needed for each phase. As a result, there is only one energy equation for both phases and heat transfer effect between different phases will be ignored.
%\item We assume immediate dynamic equilibrium so that no separate momentum equation is needed for each phase. As a result, there is only one vector moment equation for both phases. Drag force term will not show up in the momentum equation as two phases always move with the same velocity.
%\item Because of above assumptions, all other micro-physics process (like phase change of $H_2O$, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model.
%\end{itemize}
%Finally the governing equations for this model in Eulerian form are:
\begin{equation}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho}
\end{equation}
\begin{equation}
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks}
\end{equation}
\begin{equation}
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v}
\end{equation}
\begin{equation}
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{equation}
$\xi$ is the mass fraction of ejected material.
$E = e + K $ is total energy which is summation of kinetic energy $K$ and internal energy $e$.
An additional equation is required to close the system. In this model, the equation for closing the system is an EOS:
\begin{align}
p = (\gamma_m - 1)\rho e \label{eq:EOS}
\end{align}
Where 
\begin{equation}
\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
\end{equation}
\begin{equation}
Rm = n_gR_g + n_aR_a  \label{eq:gov-Rm}
\end{equation}
\begin{equation}
C_{vm} = n_s C_{vs} + n_g C_{vg} + n_a C_{va} \label{eq:gov-Cvm}
\end{equation}
\begin{equation}
n_a = 1 - \xi \label{eq:gov-na}
\end{equation}
\begin{equation}
n_g = \xi n_{g0} \label{eq:gov-ng}
\end{equation}
\begin{equation}
n_s = \xi - n_g \label{eq:gov-ns}
\end{equation}
Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscription $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\
In current model the initial domain is a 3D box. The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
At the vent, $T$, $\textbf{v}=\{0,0,150\}^T$, $p=1.01\times10^5 Pa$, $n_{g0}=0.05$ and mass discharge rate $\dot M$ is given. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the bounday. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, boundary condtions for density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Involved_initial} and figure \ref{fig:Involved_domain_adjusting}). 
\subsection{Discretized governing equations with SPH}
There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive view over SPH. We will not cover these basic theory of SPH in this paper.
The discretized governing equations with SPH are:
\begin{equation}
<\rho_a^a>=\sum m_b w_{ab} (h_a) \label{eq:gov-sph-d1}
\end{equation}
\begin{equation}
<\rho_i^{sg}>=\sum_j m_j w_{ij} (h_i) \label{eq:gov-sph-d2}
\end{equation}
\begin{equation}
\begin{split}
<\dfrac{d \textbf{v}_{\alpha}}{d t}>= &
-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}) \\ &\bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} \\ & + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+\textbf{g}
\end{split} 
\label{eq:gov-sph-v}
\end{equation}
\begin{equation}
\begin{split}
<\dfrac{d e_{\alpha}}{d t}>=&
 0.5\sum_b [m_b \textbf{v}_{\alpha b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b})\\ & \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
+0.5\sum_j [m_j \textbf{v}_{\alpha b}(\dfrac{p_j}{\rho_j^2} \\ & + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
\end{split} 
\label{eq:gov-sph-e}
\end{equation}
Where
$\rho_a^a$ is density of phase 1 (air).
$\rho_i^{sg}$ is density of phase 2 (erupted material).
$\rho=\rho^a + \rho^{sg}$ is density of mixture of phase 1 and phase 2.
\begin{equation}
\textbf{v}_{\alpha b}=\textbf{v}_{\alpha}-\textbf{v}_b
\end{equation}
\begin{equation}
\textbf{v}_{\alpha j}=\textbf{v}_{\alpha}-\textbf{v}_j
\end{equation}
$w_{ab} (h_a)= w(\textbf{r}_a-\textbf{r}_b,h_a)$ is smoothing kernel.
%$\bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})$-->need double check
$\Pi$ is artificial viscosity term \cite{monaghan1992smoothed}.
Index $a$, $b$ is for phase 1.
Index $i$, $j$ is for phase 2.
Index $\alpha$, $\beta$ can be index of either phase 1 or phase 2.
The position of each particle is updated according to the following equation.
\begin{align}
\dfrac{d \textbf{r}_a}{dt} = \textbf{v} \label{eq:gov-update-pos}
\end{align}
Free surface flow are in nature turbulent. We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume. This will result to a filtered velocity for position updating and additional turbulent terms in momentum and energy eqution.
\subsection{Solver performance}
\begin{figure}[!t]
\centering
\includegraphics[scale=0.33]{2cases_time}
\caption{Excuting time of test case 1 and test case 2}
\label{fig:2cases_time}
\end{figure}
%
\begin{figure}[!t]
\centering
\includegraphics[scale=0.33]{2cases_efficiency}
\caption{Influnece of total work load on strong scalability}
\label{fig:2cases_efficiency}
\end{figure}
%
\begin{figure}[!t]
\centering
\includegraphics[scale=0.33]{2D_vs_3D_efficiency}
\caption{Strong scalability of 3D domain decomposition and 2D domain decomposition}
\label{fig:2D_vs_3D_efficiency}
\end{figure}
Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0, 4000]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2. The computational work load of test case 2 is 8 times of that of the test case 1. The computation was run for 20s physical time. The running time and speed up are shows in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}. The results shows linear speed up when number of processes is smaller than 16. The overhead of strong scalability can be increased by increasing total amount of work load which is implied by the fact that test case2 has better speed up than test case1 when number of processes is larger than 16. We also compared the performance of 2D domain decomposition and 3D domain decomposition. As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better strong scalability than 3D domain decomposition. When number processes increase, 3D domain decomposition shows a better speed up. The total number of footprint buckets in x-y plane is 64 for the 2D and 3D domain decomposition test case. When we request 128 processes, 64 processes will be totally idle at the beginning because there is only 64 buckets initially on x-y plane. As simulation goes, the domain will expand and more and more idle processes will start getting tasks. That's why we still get some speed up when number of process is 128.
\section{Conclusion}
We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology that is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particles. 
Hashtables with external link list are adopted for accessing  particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The code was further improved to 4 times faster by adjusting computational domain according to progress of simulation. 
Performance tests on our code shew a linear speed up and an overhead cast by total among of work load.
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captions off option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
\IEEEtriggeratref{28}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
\bibliography{Reference}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2016-01-04 Paper for submission.tex
add weak scalability

@@
 %to distinguish particles added at the same position but different simulation time step. 
-Both particle and background grid are managed by hashtables which ensures quick and flexible accessing, adding and deleting of element.
+Both particle and background grid are managed by hashtables which ensure quick and flexible accessing, adding and deleting of element.
 Additional link list attached to the hashtable is used to handle hash confliction.
 A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance.
-Furtherly, computational domain is adjusted during simulation to reduce computational cost.
-%Numerical tests show that our code has linear speed up.
+In addition, computational domain is adjusted during simulation to reduce computational cost.
+Numerical tests show that our code has acceptable strong scalability and good weak scalability.
 These strategies can be further applied to many other implementations of meshfree methods, especially those implemetations that require flexible accessing, adding and deleting of particles and backgroud grids.
@@
 In this paper, we implement SPH to simulate a simplified version of complicated volcanic plume: the JPUE, and develop data management schemes for it.\\
-Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing.
+Among existing CPU parallel SPH schemes, most of them focus on neighbors searching algorithm and dynamic load balancing.
 (eg.\cite{ferrari2009new, crespo2015dualsphysics}).
@@
 We define C++ classes which contains all data of particle and bucket.
-As for the storage of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
+As for the management of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
 Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier.
-The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have a different identifier.
+The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
 As for domain decomposion, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming.
@@
 \includegraphics[scale=0.34]{Work_flow}
-\caption{Basic Work flow for SPH}
+\caption{Basic work flow for SPH}
 \label{fig:Work_flow}
@@
 New particles also need to be added for the eruption boundary condition.
-We will describe in this section these strategies which allows flexible deleting and adding.
+We will describe in this section these strategies which satisfies these demand.
 \subsection{SFC based indexing}
@@
 \subsubsection{Particle and bucket}
-The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbor searching and domain decomposition.
+The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbors searching and domain decomposition.
 Both are defined as classes in C++.
-Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, like density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg.
-pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is vector of particle keys in our application).
-Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (27 neighbor buckets including itself for three dimension and 9 for two dimension) and possessed particles.
+Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, eg.
+density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg.
+pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is a vector of particle keys in our application).
+Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and possessed particles.
 Objects defined based on these two classes are then accesse through hash tables.
@@
 How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table.
-As the distribution of keys is determined by particles' initial location (or centroids of buckets), the hash table size is under our control.
+As the distribution of keys is determined by particles' initial locations (or centroids of buckets), the hash table size is under our control.
 We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency.
 Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions.
-Abani\cite{patra2003data} did numerical experiments to make trade-off to determine hash table size.\\
+Abani\cite{patra2003data} did numerical experiments to examine
+the effect of the table size on the dierent data management operations.\\
  %We should also do some experiment to optimize our hash table size.
@@
 \includegraphics[scale=0.255]{Particle_adding_with_link}
-\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, these new particles will be stored in external link list}
+\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, pointer to these new particles will be stored in external link list.}
 \label{fig:Particle_adding_with_link}
@@
 \end{equation}
-One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as higher digit and $t$ as lower digit of the large number.
+One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as the higher digit and $t$ as the lower digit of the large number.
 However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area.
@@
 To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}).
-\subsection{Load balance strategy}
+\subsection{Load balancing strategy}
 \subsubsection{Weighted work load}
 Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle.
-Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}).
+Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial}).
 As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data.
@@
 The SFC sequence passing through centroids of all buckets now becomes a weighted sequence.
-Domain decomposition will conducted based on the weighted SFC of buckets.
+Domain decomposition will conducted based on the weighted SFC of buckets.\\
 \begin{table}[t!]
@@
   \centering
-  \caption{Computational Cost for Different Steps Per Particle}
+  \caption{Computational Cost Per Particle for Different Steps}
   \label{tab:Computational_cost_steps}
@@
 A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}.
-A scalability comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger.
+A comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger.
 %Some parallel implementations of SPH use 2D decomposition \cite{kumar2013parallel} even 1D decomposition \cite{marrone2012study, cherfils2012josephine}.  
@@
 Movement of particles, adding of new particles, adjusting of domain will lead to important load imbalance between processes.
-To handle this, computational load is monitored at given interval (The interval is optimized based on numerical experiments).
-Repartitioning is carried out if load imbalance is larger than a given tolenrance.
+To handle this, computational load is monitored at a given interval (The interval is optimized based on numerical experiments).
+And repartitioning is carried out when load imbalance is larger than a given tolerance.
 \\
-%Figure \ref{fig:t0_domain_decomposition} and \ref{fig:t10_domain_decomposition} show the initial decomposition and domain decomposion at 10 seconds.\\
-
-%We adopt the same SFC based partition strategy as described by Patra \ref{patra1999efficient}.
-
-%\begin{figure}[h]
-%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: initial domain decomposition}
-%\centering
-%\label{fig:t0_domain_decomposition}
-%\includegraphics[width=0.5\textwidth]{t0_domain_decomposition}
-%\end{figure}
-%\begin{figure}[h]
-%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: decomposition at t=10s}
-%\centering
-%\label{fig:t10_domain_decomposition}
-%\includegraphics[width=0.5\textwidth]{t10_domain_decomposition}
-%\end{figure}
 As some of the neighbor particles reside in other
@@
 However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone.
-Because the whole domain, which occupied by stationary fluid before eject fluids reaching there, has to be discretized at the very beginning of simulation.
+Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation.
 A lot of CPU time will be spent on computing of "stationary" particles.
@@
 The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\
-\begin{figure}[h]
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.33]{Work_flow_adjust}
 \caption{Work flow that enables domain adjusting feature}
-\centering
 \label{fig:Work_flow_adjust}
-\includegraphics[width=0.5\textwidth]{Work_flow_adjust}
 \end{figure}
@@
 The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles.
-So they will be involved in the mixing process in the near future.
+So they will be involved in the near future.
 These not involved particles are particles that far away from the eject source and are still at the initial state.
@@
 As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced.
-When originally stationary air is influenced by erupted material, the mass fraction of the erupted material of mixture will increase from zero to a positive value.
-So we can determine whether a particle is involved or not is based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation).
+When originally stationary air is influenced by erupted material, the mass fraction of the erupted material will increase from zero to a positive value.
+So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation).
 Other physical properties, such velocity, can also serve as alternative "swich criteria".\\
@@
 A has-involved-flag is added to bucket class, too.
-Buckets are also categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0).
+Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0).
 Bucket that has any involved particle will be set to be has involved (has-involved-flag=3).
@@
 New pressure ghost particles and wall ghost particles will be added around the adjusted domain.
-This domain adjusting process is shown in figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}.
+This domain adjusting process is shown in figure \ref{fig:Domain_adjusting}.
 The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}\\
@@
 \subsection{Governing equations\cite{suzuki2005numerical} and boundary conditions}
-Based on Navier-Stokes equations and several simplification assumption, the governing equations in Eulerian form are:
+Based on Navier-Stokes equations and several simplifications, the governing equations in Eulerian form are:
 %the following assumptions are made:
@@
 The pressure of the surrounding atmosphere is specified on pressure boundaries.
-Except for the pressure, boundary condtions for density, velocity, and energy will depend on the solution.
-We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Involved_initial} and figure \ref{fig:Involved_domain_adjusting}).
+Except for the pressure, density, velocity, and energy will depend on the solution.
+We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}).
 \subsection{Discretized governing equations with SPH}
@@
 \end{figure}
-Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo.
-The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0, 4000]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2.
+%
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.33]{weak_loglog}
+\caption{Weak scalability test results}
+\label{fig:weak_loglog}
+\end{figure}
+%Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
+The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2.
 The computational work load of test case 2 is 8 times of that of the test case 1.
-The computation was run for 20s physical time.
-The running time and speed up are shows in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}.
-The results shows linear speed up when number of processes is smaller than 16.
-The overhead of strong scalability can be increased by increasing total amount of work load which is implied by the fact that test case2 has better speed up than test case1 when number of processes is larger than 16.
+The simulations run for 20s physical time.
+The consumed time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}.
+Linear speed up is observed when number of processes is smaller than 16.
+Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load.
 We also compared the performance of 2D domain decomposition and 3D domain decomposition.
-As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better strong scalability than 3D domain decomposition.
-When number processes increase, 3D domain decomposition shows a better speed up.
-The total number of footprint buckets in x-y plane is 64 for the 2D and 3D domain decomposition test case.
-When we request 128 processes, 64 processes will be totally idle at the beginning because there is only 64 buckets initially on x-y plane.
+As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processese.
+This can be explained by the fact that 2D domian decomposition can get more regular subdomain and as a results, need less communication.
+When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.
+The weak scalability test is conducted with the same initial domain and various smoothing length.
+Each simulation runs for 400 loops.
+The average number of real particles of each process keep constant.
+As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132.
+the relative sharp increase in simulation time from 32 processes to 132 processes might cause by a increase of number of neighbor processes.
+%That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
+%When we request 128 processes, 64 processes will be totally idle at the beginning.
 As simulation goes, the domain will expand and more and more idle processes will start getting tasks.
 That's why we still get some speed up when number of process is 128.
+3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.
 \section{Conclusion}
@@
 Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation.
-SFC based index scheme, which provides a global numberring methodology that is purely coordinates dependent, is adopted to give each bucket an unique identifier.
-A time dependent key which is also based on SFC is used as identifier for particles.
+SFC based index scheme, which provides a global numberring methodology which is purely coordinates dependent, is adopted to give each bucket an unique identifier.
+A time dependent key which is also based on SFC is used as identifier for particle.
 Hashtables with external link list are adopted for accessing particles and buckets data.
 Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval.
-The code was further improved to 4 times faster by adjusting computational domain according to progress of simulation.
-Performance tests on our code shew a linear speed up and an overhead cast by total among of work load.
+The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation.
+Scalability tests on our code shew acceptable strong scalability and good weak scalability.
+3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.
 % An example of a floating figure using the graphicx package.
@@
 
-The authors would like to thank...
+Computational results reported here were performed at
+the Center for Computational Research at the University at
+Buffalo. This project is supported by Grants No. NSF
+1131074 from the National Science
+Foundation.
 

### 2016-01-06 Paper for submission.tex
fixed 1 2.1

@@
 % affiliations
-\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra}
+\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra, Matt Jones, ...}
 \IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\
@@
 Both particle and background grid are managed by hashtables which ensure quick and flexible accessing, adding and deleting of element.
-Additional link list attached to the hashtable is used to handle hash confliction.
+%Additional link list attached to the hashtable is used to handle hash conflicts.
  A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance.
@@
 Numerical tests show that our code has acceptable strong scalability and good weak scalability.
-These strategies can be further applied to many other implementations of meshfree methods, especially those implemetations that require flexible accessing, adding and deleting of particles and backgroud grids.
+These strategies can be further applied to many other implementations of meshfree methods, especially those implementations that require irregular
+particle distributions, adding and deleting of particles and leverage adaptive background grids.
 \end{abstract}
@@
 % no \IEEEPARstart
-SPH is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
+SPH (spherical particle hydrodynamics) is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
 Subsequently it was extended to large strain solid mechanics and computational fluid mechanics. 
-A lot of research has been done in parallellization of SPH. 
+A lot of research has appeared  on  parallellization of SPH in the last few years. 
 Goozee\cite{goozee2003distributed} implemented a simple SPH code using MPI, OpenMP and BSP. 
-Chen Wenbo\cite{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
-David W. Holmes\cite{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
-Domínguez\cite{dominguez2011optimization} presented optimizations for both CPU and GPU parallelization of a SPH method. 
-Angela Ferrari\cite{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme.
-D. Kumar and Abani\cite{kumar2013parallel} implemented CPU parallel Godunov SPH in simulation of granular flows.
-A.J.C. Crespo\cite{crespo2015dualsphysics} used the parallel power computing of GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
-However, most implementation of parallel SPH method is limitted to bechmark problems like dam break, or relative simple scenarios like breaking-waves, floodings ect.
-Rare work has been done in simulation of more complicated problems, such as eruption of volcano plume which is essentially a mutilple phases, turbulent, ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction ect.
-Prediction of such more complicated phenomena with acceptable accuracy at given time window can not been acomplished without parallel computing.
-What's more, imposing of some types of boundary conditions (such as realistic wind feild, eruption boundary condition) requires dynamiclly adding and removing of particles during simulation.
-This requies efficient and more flexible data management scheme.
-What need to point out is that the benefit of flexible data management strategies are not only limited to specific implementations of SPH.
-Actually, flexibility in data management are more desperately needed in some advanced techniques of SPH, such as dynamic particle splitting techniques\cite{vacondio2012accurate, feldman2007dynamic}, which will give greater resolution at the area of interest by spliiting one large particle to several smaller ones.
+ Wenbo\cite{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
+  Holmes\cite{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
+Dominguez\cite{dominguez2011optimization} presented optimizations for both CPU and GPU parallelization of a SPH method. 
+ Ferrari\cite{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme.
+Kumar et. al. \cite{kumar2013parallel} implemented a parallel Godunov SPH in simulation of granular flows.
+ Crespo\cite{crespo2015dualsphysics} used the GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
+However, most implementation of parallel SPH method presented to date are limited to standard SPH and benchmark problems like dam break, or relatively
+ simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume which is essentially a mult-phase turbulent flow,
+  ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction etc. is relatively rare. 
+  Prediction of such  c complicated phenomena with acceptable accuracy at given time window requires resolutions (very high particle counts) that cannot been accomplished without parallel computing.
+What's more, imposing of some types of boundary conditions (such as realistic wind field, eruption boundary condition) requires dynamically adding and removing of particles during simulation.
+This requires efficient and more flexible data management scheme.
+What we need to point out is that the benefit of flexible data management strategies are not only limited to specific implementations of SPH.
+Actually, flexibility in data management are more critical for several advanced techniques of SPH, such as dynamic particle splitting techniques\cite{vacondio2012accurate, feldman2007dynamic}, which will give greater resolution at the area of interest by splitting one large particle to several smaller ones.
 In this paper, we implement SPH to simulate a simplified version of complicated volcanic plume: the JPUE, and develop data management schemes for it.\\
@@
 Less attention has been paid to developing of more flexible data management schemes for more complicated problems.
-Fortunately, efficient and flexible data management strategies for high performance computing have been successfully implemented in mesh based methods(eg.
+Fortunately, efficient and flexible data management strategies for high performance computing have been successfully implemented in mesh based methods (eg.
 \cite{laszloffy2000simple} for adaptive hp FEM, and \cite{pitman2003computing, patra2005parallel} for FVM).
@@
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-We adopted background grid which was proposed by Monaghan and Lattanzio\cite{monaghan1985refined} and implemented widespreadly in parallel SPH.
+Of the many choices possible for this we adopted a background grid which was proposed by Monaghan and Lattanzio\cite{monaghan1985refined} and is quite popular in parallel SPH.
 The background grid is also used for domain decomposition in SPH.
-We refer to the elements of background grid, namely a squares for two dimension and cubics for three dimension, as buckets.
-As for the storage of physical quantities associated to each particle, different strategies was adopted in existing implementations of SPH.
+We refer to the elements of background grid, namely  squares for two dimension and cubics for three dimension, as buckets.
+
+As for the actual storage of data representing the physical quantities associated to each particle, different strategies  have been adopted in existing implementations of SPH. 
 In both SPHysics\cite{dominguez2011optimization} and DualSPHysics\cite{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
-This has two advantages: 1)access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
+This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
 But adding, deleting and especially accessing of particles are cumbersome.
-Angela Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
+Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
 Storage problems caused by fix-size arrays are thereby also eliminated.
 We define C++ classes which contains all data of particle and bucket.
-As for the management of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
-Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier.
-The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
-As for domain decomposion, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming.
+As for the management of data, we adopt a hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
+Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier -- a strategy known to preserve some locality at minimal cost.
+The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
+
+As for domain decomposition, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
 So we adopt an easy-programming scheme based on SFC \cite{patra1999efficient}.\\
-To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation.
+To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation needs.
 For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluid.
-This feature is accomplished by adding a scan function to monitor the most outside layer of the domain and turn ghost particles to real particles at proper time.
+This feature is accomplished by adding a scan function to monitor the outermost layer of the domain and turn ghost particles to real particles at the proper time.
 %The computational cost of this additional scan function is low.
@@
 SPH is a meshfree, Lagrangian method.
-The domain is discretized by particles and the position of particle is updated at every time step.
-The physical laws (such as conservation laws of mass, momentum and energy) written in the form of partial differential equations need to be transformed into the Lagrangian formalism.
-Using a kernel function that provides the weighted estimate of the field variables at a discrete point (particle), the integral equations are evaluated as sums over neighbor particles.
+The domain is discretized by particles and the position of each particle is updated at every time step.
+The physical laws (such as conservation laws of mass, momentum and energy) written in the form of partial differential equations need to be transformed into the Lagrangian particle formalism of SPH.
+Using a kernel function that provides the weighted estimate of the field variables in the neighborhood of a discrete point (particle), the integral equations are evaluated as sums over neighbor particles.
 Only particles located within support of kernel function will interact.
-Thus, physical proterties (position, density, velocity, internal energy, pressure) are updated based on its neighbors.
-So a neighbor searching need to be carried out before updating of physical proterties.
-We use buckets which overlap the whole domain and keep fixed in time during the entire simulation, to reduce searching cost.
-Domain decomposition will be based on SFC going through centroids of all buckets.
-A basic works flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
+Thus, physical properties (position, density, velocity, internal energy, pressure) are updated based on its neighbors.
+So a neighbor search needs to be carried out before updating of physical properties.
+We use buckets which contain all particles associated with a sub-domain
+ and are kept fixed in time during the entire simulation, to reduce search cost (since search can now be restricted to only neighboring domains).
+  Domain decomposition will be based on an SFC going through centroids of all buckets. A basic work flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
 Particles need to be added and(or) removed during simulation of JPUE problems.
@@
 New particles also need to be added for the eruption boundary condition.
-We will describe in this section these strategies which satisfies these demand.
+We will describe in this section   strategies which satisfies these requirements.
 \subsection{SFC based indexing}
@@
 For example, new particles need to be added at the bottom of the eject vent for JPUE simulation.
-To distinguish particles added at the "same place" (the small area, all points whithin which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
+To distinguish particles added at the ``same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
 The time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step.
@@
 Since SPH particles only interact with its neighbors, geometric locality can be exploited for efficient storage and retrieval of bucket and particle data.
-\item This type of key effectivelly generate a global address space.
-Globality of key and conservation of locality make it easy to partition the sorted key squence and obtain the decompositiion of the problem.
+\item This type of key effectively generates a global address space.
+Globality of key and conservation of locality make it easy to partition the sorted key sequence and obtain a decompositiion of the problem.
 \end{itemize}
-What need emphasize here is that motion of particles might mess up locality that established based on initial coordinates of particles.
-As particles are moving regularly, the locality of most of particles should still be conserved during simulation.
+What we need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles.
+However, as particles are moving fairly regularly in a JPUE, the locality of most of particles should still be conserved during simulation.
 %If we can design a re-numberring method which is computationally cheap, we can restore the locality.

### 2016-01-06 Paper for submission.tex
fixed all sections

@@
 \centering
-\subfloat[SFCs passing all particles and buckets]{\includegraphics[scale=0.20]{SFC_particles_buckets}}
+\subfloat[SFCs passing all particles and buckets]{\includegraphics[scale=0.15]{SFC_particles_buckets}}
 \label{fig:SFC_particles_buckets}
 \hfil
-\subfloat[Domain decomposition based on SFC of buckets]{\includegraphics[scale = 0.22]{SFC_particles_buckets_partition}}
+\subfloat[Domain decomposition based on SFC of buckets]{\includegraphics[scale = 0.16]{SFC_particles_buckets_partition}}
 \label{fig:SFC_particles_buckets_partition}
@@
 \end{itemize}
-What we need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles.
-However, as particles are moving fairly regularly in a JPUE, the locality of most of particles should still be conserved during simulation.
-%If we can design a re-numberring method which is computationally cheap, we can restore the locality.
+We need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles.
+However, as particles are moving fairly regularly in a JPUE, the locality of most of particles is largely conserved during simulation.
+We will revisit this question for more chaotic flows and
+ design a re-numbering method which is computationally cheap, to restore the locality.
 \subsection{Data structure}
 \subsubsection{Particle and bucket}
-The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbors searching and domain decomposition.
+The most basic data structure of SPH are particle, for problem description, and bucket, for neighbors searches and domain decomposition.
 Both are defined as classes in C++.
-Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, eg.
-density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg.
-pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is a vector of particle keys in our application).
-Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and possessed particles.
-Objects defined based on these two classes are then accesse through hash tables.
+Information that is contained in particle class can be categorized into six categories: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknowns in governing equations, e.g.
+density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeated computations eg.
+pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases  ) and neighbor information (it is a vector of particle keys in our application).
+Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and owned particles.
+Objects defined based on these two classes are then accessed through hash tables.
 \subsubsection{Hash table and hash confliction}
 As discussed at the beginning of this section, implementation of SPH in more realistic scenarios requires dynamic memory management and flexible data access.
-One of the fundemental data structures that satisfy such requirement is hash table.
+One of the fundamental data structures that satisfy such requirement is hash table.
 Another option is B-tree.
 We adopt hash table.
-An implementation of B-tree under a similar situation for mesh based methods can be found in other papers(eg.\cite{patra2003data}).\\
+An implementation of B-tree under a similar situation for mesh based methods can be found in other papers (eg.\cite{patra2003data}).\\
 Hash table, which is divided into slots, are array based data structure.
@@
 \end{equation}
-The hash table has O(1) data accessing, adding and deleting properties when there is no confliction.
-How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table.
+The hash table has O(1) data access, adding and deleting properties when there is no conflict.
+How many conflicts will happen depends on both distribution of keys in the key space and size of the hash table.
 As the distribution of keys is determined by particles' initial locations (or centroids of buckets), the hash table size is under our control.
-We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency.
-Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions.
-Abani\cite{patra2003data} did numerical experiments to examine
-the effect of the table size on the dierent data management operations.\\
+We can use a very large hash table to minimize hash conflicts at the expense of sparse data distribution which will lead to high cache misses and low memory efficiency.
+Or alternately, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflicts.
+Patra et al.\cite{patra2003data} did numerical experiments to examine
+the effect of the table size on the different data management operations.\\
  %We should also do some experiment to optimize our hash table size.
-One way to handle hash confliction is using an additional sorted vector attached to the hash table.
+One way to handle hash conflicts is by using an additional sorted vector attached to the hash table.
 When several keys hash to the same slot, a vector will be created.
-The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrievaling .
-Another option to handle hash confliction is using an additional link list which is more flexible in memory allocation.
+The vector is sorted based on keys so that a binary search can be used to find the correct position for adding, deleting or retrieving.
+Another option to handle hash conflicts is using an additional link list which is more flexible in memory allocation.
 The average time complexity of binary search is O(log n) while that for linear search based on link list is O(n).
-However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longger.
-Choosing of proper way to handle hash confliction greatly depends on the problem itself.
-For the test problem in this paper, successively adding of particles at the bottom of the eject vent will lead to hash confliction of many particles, which implies a long link list.
-But considering the very long conflictions only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash confliction.
+However, accessing efficiency of link list is much lower than array based data structure, especially when the link list becomes longer.
+Choosing of proper way to handle hash conflicts greatly depends on the problem itself.
+For the test problem in this paper, successively adding particles at the bottom of the eject vent will lead to hash conflicts of many particles, which implies a long link list.
+But considering the very long conflicts only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash conflicts.
 This decision was made based on numerical experiments.\\
@@
 \includegraphics[scale=0.255]{Particle_adding_with_link}
-\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, pointer to these new particles will be stored in external link list.}
+\caption{Non-uniform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portion of the domain, pointer to these new particles will be stored in external link list.}
 \label{fig:Particle_adding_with_link}
@@
 However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area.
-Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent.
-That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}.
-To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}).
+Only ghost particles for eruption boundary condition will be successively added at the same place: bottom of the vent.
+That is to say, particles are distributed non-uniformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}.
+To avoid non-uniform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}).
 \subsection{Load balancing strategy}
@@
 The SFC sequence passing through centroids of all buckets now becomes a weighted sequence.
-Domain decomposition will conducted based on the weighted SFC of buckets.\\
+Domain decomposition  is conducted based on the weighted SFC of buckets.\\
 \begin{table}[t!]
@@
 \subsubsection{Domain decomposition and dynamic load balancing}
-Domain decomposition will be conducted based the weighted SFC of buckets.
+%Domain decomposition will be conducted based the weighted SFC of buckets.
  Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets.
@@
 As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition.
-A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}.
-A comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger.
+A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Kumar et al \cite{kumar2013parallel}.
+A comparison of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger.
 %Some parallel implementations of SPH use 2D decomposition \cite{kumar2013parallel} even 1D decomposition \cite{marrone2012study, cherfils2012josephine}.  
@@
 As some of the neighbor particles reside in other
-partitions.
-A set of “guest” particles and buckets are used to synchronize data across partitions.
-To minimize communications, data is synchronized only where needed, using non-blocking MPI communications.
+partitions. A set of  particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 
 \section{Adjusting of Domain During Simulation}
 %SPH particles can be viewed as discretization points\cite{diltsg1999moving}. 
-As a Lagrangian method, SPH is able to automatically adjust computational domain as the postition of the dicretization points are updated at every time step.
-However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone.
-Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation.
-A lot of CPU time will be spent on computing of "stationary" particles.
-It is pure wasting of computational resources.
-If simulating of stationary particls can be avoided, the computational cost will be reduced greatly.
-To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation.
+As a Lagrangian method, SPH is able to automatically define and adjust its computational domain as the position of the discretization points are updated at every time step.  The 
+collective support of the SPH particles defines the computational domain.
+However, for JPUE simulation (and many similar/related phenomena), where some fluid ejects into stationary fluid and gets mixed, the domain-adjusting feature of SPH cannot be taken advantage of as all of the stationary fluid must be represented.
+%Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation.
+A lot of CPU time will be spent on computing associated with these stationary particles.
+If simulating of stationary particles can be avoided, the computational cost will be reduced greatly.
+%To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation.
 We propose a simple strategy to add such feature in our code with low computational cost.
@@
 When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition.
-The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\
+The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such a feature (see figure \ref{fig:Work_flow_adjust}).\\
 \begin{figure}[!t]
@@
 \end{figure}
-The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class.
-Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0).
-The involved particles are particles that have already been involved in eject mixing.
-The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles.
-So they will be involved in the near future.
-These not involved particles are particles that far away from the eject source and are still at the initial state.
+This is implemented by using a  involved-flag, in the particle class.
+Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potentially involved (involved-flag=1), and not involved (involved-flag=0).
+The involved particles are particles that have already been affected by the mixing.
+The potentially involved particles are particles that have not been involved in mixing but are adjacent to involved particles and will thus be be involved in the near future.
+%These not involved particles are particles that far away from the eject source and are still at the initial state.
 For these not involved particles, it is meaningless to update its physical properties.
 As a consequence, searching for neighbors and data communitation for these particles are also unnecessary.
-That is to say, only potential involved and involved particles need to be simulated.
-As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced.
+All communication and computation associated with uninvolved particles can be ignored. That is to say, only potential involved and involved particles need to be simulated.
+As simulation progresses, the ejected fluid will reach larger area and more and more particles will be influenced.
 When originally stationary air is influenced by erupted material, the mass fraction of the erupted material will increase from zero to a positive value.
-So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation).
-Other physical properties, such velocity, can also serve as alternative "swich criteria".\\
+So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold ($10^{-5} $ in our simulation).
+Other physical properties, such velocity, can also serve as alternative ``swich criteria".\\
 \begin{figure*}[!t]
@@
 \label{fig:Involved_domain_adjusting}
-\caption{Domain adjusting based on Involved flag of particles}
+\caption{ Adjusting the domain based on Involved flag of particles}
 \label{fig:Domain_adjusting}
 \end{figure*}
-A has-involved-flag is added to bucket class, too.
-Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0).
+A similar scheme can be deployed for buckets too resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
+%A has-involved-flag is added to bucket class, too.
+Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has not involved (has-involved-flag=0).
 Bucket that has any involved particle will be set to be has involved (has-involved-flag=3).
 And its neighbor buckets that do not contain any involved particle will be set to be has potential involved.
-All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles.
+%All particles in has involved buckets or has potential involved buckets will be set to be potential involved except for involved particles.
 There are two situations that will switch a has potential involved bucket to be a has involved bucket.
@@
 Second, when mass fraction of a particle (It should be a potental involved particle) exceeds the threshold, this particle will turn to a involved particles and if this is the first involved particle in the bucket, the bucket will turn from a has potential involved bucket to a has involved bucekt.\\
-The most outside buckets layer of all has potential involved buckets will be scaned every time after updating.
-If any of these buckets becomes has involved, the domain will be enlarged by turnning the pressure ghost particles to real particles and switching involved-flag from 0 to 1.
+%The most outside buckets layer of all has potential involved buckets will be scaned every time after updating.
+If any of these buckets becomes has involved, the domain will be enlarged by turning the pressure ghost particles to real particles and switching involved-flag from 0 to 1.
 Also, the buckets that originally contain pressure ghost particles will become has potential involved.
@@
 The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}\\
-For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption.
-When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$.
-This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
-Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
+
 \section{Numerical Test}
-As we are targeting at developing data management and paralell strategies for more complicated implementations of SPH which demmand quick and flexible data access, delete and add.
-The test problem should have such demand.
-Volcanic eruption which is essentially a mutilple phases, turbulent, ejection mixing flow accompanied with microphysics processes requires more flexibile data management.
-We adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem.
+
+Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
+Thus, the test problem used should have such requirements.
+Simulations of a volcanic eruption, which is essentially a muti-phase, turbulent, ejection mixing flow accompanied with microphysics processes will place these demands on the framework.
+Thus, we adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem.
 In this model, one phase is air while another phase is ejected material.
@@
 %Finally the governing equations for this model in Eulerian form are:
-\begin{equation}
-\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho}
-\end{equation}
-\begin{equation}
-\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks}
-\end{equation}
-\begin{equation}
-\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v}
-\end{equation}
-\begin{equation}
-\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
-\end{equation}
+\begin{eqnarray}
+\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
+%\end{equation}
+%\begin{equation}
+\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) &=& 0 \label{eq:gov-cs-ks}\\
+%\end{equation}
+%\begin{equation}
+\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) &=& \rho \textbf{g} \label{eq:gov-cs-v} \\
+%\end{equation}
+%\begin{equation}
+\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] &=& \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
+\end{eqnarray}
 $\xi$ is the mass fraction of ejected material.
@@
 \end{equation}
-Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant.
-The subscription $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\
+Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscript
+ $m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\
 In current model the initial domain is a 3D box.
@@
 Velocity is zero for non-slip wall boundary.
-We assume the boundary to be adiabatic and the heat flux is zero on the bounday.
+We assume the boundary to be adiabatic and the heat flux is zero on the boundary.
 The pressure of the surrounding atmosphere is specified on pressure boundaries.
@@
 \subsection{Discretized governing equations with SPH}
-There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive view over SPH.
-We will not cover these basic theory of SPH in this paper.
+There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.
+These  may be consulted for additional detail.%We will not cover the basic theory of SPH in this paper.
 The discretized governing equations with SPH are:
@@
 \end{align}
-Free surface flow are in nature turbulent.
+Free surface flows are inherently turbulent.
 We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume.
-This will result to a filtered velocity for position updating and additional turbulent terms in momentum and energy eqution.
+This will result in a filtered velocity for position updates and additional turbulent terms in momentum and energy equation.
 \subsection{Solver performance}
@@
 \includegraphics[scale=0.33]{2cases_time}
-\caption{Excuting time of test case 1 and test case 2}
+\caption{Execution time of test case 1 and test case 2}
 \label{fig:2cases_time}
@@
 \includegraphics[scale=0.33]{2cases_efficiency}
-\caption{Influnece of total work load on strong scalability}
+\caption{Influence of total work load on strong scalability}
 \label{fig:2cases_efficiency}
@@
 \end{figure}
-%Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
-The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2.
+Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. Two types of processors -- 8 core XXXXX running at YYY
+clock speed with ZZZ memory per core and 12 core AAAAA running at BBBB
+clock speed with CCC memory per core. Each node comprised of two sockets with two of these processors. Memory and level 3 cache are shared on each node.
+An important caveat in interpreting these results is that they were obtained in settings where the processors were obtained under normal scheduling (hence readily available) wherein the nodes may be shared with other users.
+Thus our results may suffer from  contention for memory, CPU-memory bandwidth and network access.
+As of this writing, we are waiting for exclusive access which will undoubtedly provide better results.
+On the other hand, these results are representative of what will be available during production use.
+
+The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively.
 The computational work load of test case 2 is 8 times of that of the test case 1.
 The simulations run for 20s physical time.
-The consumed time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}.
+Wall clock time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}.
 Linear speed up is observed when number of processes is smaller than 16.
@@
 We also compared the performance of 2D domain decomposition and 3D domain decomposition.
-As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processese.
-This can be explained by the fact that 2D domian decomposition can get more regular subdomain and as a results, need less communication.
+As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processes.
+This can be explained by the fact that 2D domain decomposition can get more regular subdomain and as a results, need less communication.
 When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.
@@
 Each simulation runs for 400 loops.
-The average number of real particles of each process keep constant.
+The average number of real particles of each process keep constant at $259000$.
 As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132.
-the relative sharp increase in simulation time from 32 processes to 132 processes might cause by a increase of number of neighbor processes.
+the relative sharp increase in simulation time from 32 processes to 132 processes might be caused by an increase of number of neighbor processes.
 %That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
@@
 3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.
+
+For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption.
+When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$.
+This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
+Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
+
 \section{Conclusion}

### 2016-01-06 Paper for submission.tex
Wensday night

@@
 We refer to the elements of background grid, namely  squares for two dimension and cubics for three dimension, as buckets.
-
+%
+%
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies  have been adopted in existing implementations of SPH. 
@@
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
-
+%
 As for domain decomposition, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
@@
 New particles also need to be added for the eruption boundary condition.
-We will describe in this section   strategies which satisfies these requirements.
+We will describe in this section strategies which satisfy these requirements.
 \subsection{SFC based indexing}
@@
 \subsubsection{Weighted work load}
-Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle.
-Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial}).
+Particles used in the test problem can be categorized into four types based on the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle.
+Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Domain_adjusting}).
 As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data.
@@
 So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold ($10^{-5} $ in our simulation).
-Other physical properties, such velocity, can also serve as alternative ``swich criteria".\\
+Other physical properties, like velocity, can also serve as alternative "swich criteria".\\
 \begin{figure*}[!t]
@@
 \label{fig:Involved_domain_adjusting}
-\caption{ Adjusting the domain based on Involved flag of particles}
+\caption{ Adjusting the domain based on involved-flag of particles}
 \label{fig:Domain_adjusting}
 \end{figure*}
-A similar scheme can be deployed for buckets too resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
+A similar scheme can also be deployed for buckets resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
 %A has-involved-flag is added to bucket class, too.
@@
 The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}\\
-
+%
 \section{Numerical Test}
-
+%
+%
 Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
@@
 %Finally the governing equations for this model in Eulerian form are:
-\begin{eqnarray}
+\begin{eqnarray} %%I did not notice that I can add equation like this, useful, haha
 \dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
@@
 \centering
-\includegraphics[scale=0.33]{weak_loglog}
+\includegraphics[scale=0.35]{weak_loglog}
 \caption{Weak scalability test results}
@@
 \end{figure}
-Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. Two types of processors -- 8 core XXXXX running at YYY
-clock speed with ZZZ memory per core and 12 core AAAAA running at BBBB
-clock speed with CCC memory per core. Each node comprised of two sockets with two of these processors. Memory and level 3 cache are shared on each node.
+Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo.
+Two types of processors -- 8 core L5630 running at 2.13GHz clock rate with 3GB memory per core on a Mellanox network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic network.
+Each node comprised of two sockets with two of these processors.
+Memory and level 3 cache are shared on each node.
 An important caveat in interpreting these results is that they were obtained in settings where the processors were obtained under normal scheduling (hence readily available) wherein the nodes may be shared with other users.
@@
 On the other hand, these results are representative of what will be available during production use.
-
+%
 The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively.
@@
 Each simulation runs for 400 loops.
-The average number of real particles of each process keep constant at $259000$.
-As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132.
-the relative sharp increase in simulation time from 32 processes to 132 processes might be caused by an increase of number of neighbor processes.
+The average number of real particles of each process keep constant at $25900$.
+As shown in figure \ref{fig:weak_loglog}, simulation time are almost constant with some minor fluctutations around the average value.
+These fluctuations are due to sharing nodes with other users.
+In addition, difference in simulation time for different processors are obvious.
 %That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
@@
 3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.
-
+%
+%
 For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption.
@@
 Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
-
+%
+%
 \section{Conclusion}

### 2016-01-07 Paper for submission.tex
added plume figure

@@
 % affiliations
-\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra, Matt Jones, ...}
-\IEEEauthorblockA{Department of Mechanical and Aerospace Engineering\\
+\author{\IEEEauthorblockN{Zhixuan Cao, Abani Patra}
+\IEEEauthorblockA{Department of MAE \\%Mechanical and Aerospace Engineering
  The University at Buffalo\\
@@
 Email: zhixuanc@buffalo.edu, abani@buffalo.edu}
+\and
+\IEEEauthorblockN{Matthew Jones, Jaroslaw Zola}
+\IEEEauthorblockA{Center for Computational Research\\
+Department of CSE \\%Computer Science and Engineering
+The University at Buffalo\\
+Buffalo, New York 14260\\
+Email: jonesm@buffalo.edu, \\ jzola@buffalo.edu}
 %\and
-%\IEEEauthorblockN{Homer Simpson}
-%\IEEEauthorblockA{Twentieth Century Fox\\
-%Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\and
-%\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-%\IEEEauthorblockA{Starfleet Academy\\
-%San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212\\
-%Fax: (888) 555--1212}
+%\IEEEauthorblockN{Jaroslaw Zola}
+%\IEEEauthorblockA{The University at Buffalo\\
+%Department of Computer Science and Engineering\\
+% The University at Buffalo\\
+% Buffalo, New York 14260\\
+%Email:jzola@buffalo.edu}
 }
@@
 This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE).
-Background grid is used to reduce neighbors searching cost and decompose (redecompose) the domain.
+Background grid is used to reduce neighbors searching cost and to decompose (redecompose) the domain.
 Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid.
@@
 % no \IEEEPARstart
-SPH (spherical particle hydrodynamics) is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
-Subsequently it was extended to large strain solid mechanics and computational fluid mechanics. 
+SPH is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
+Subsequently it was extended to computational fluid mechanics. 
 A lot of research has appeared  on  parallellization of SPH in the last few years. 
@@
 However, most implementation of parallel SPH method presented to date are limited to standard SPH and benchmark problems like dam break, or relatively
- simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume which is essentially a mult-phase turbulent flow,
+simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume (figure \ref{fig:Plume}) which is essentially a mult-phase turbulent flow,
 ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction etc. is relatively rare. 
-  Prediction of such  c complicated phenomena with acceptable accuracy at given time window requires resolutions (very high particle counts) that cannot been accomplished without parallel computing.
+Prediction of such complicated phenomena with acceptable accuracy at given time window requires resolutions (very high particle counts) that cannot been accomplished without parallel computing.
 What's more, imposing of some types of boundary conditions (such as realistic wind field, eruption boundary condition) requires dynamically adding and removing of particles during simulation.
@@
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
-%
 As for domain decomposition, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
@@
 For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluid.
-This feature is accomplished by adding a scan function to monitor the outermost layer of the domain and turn ghost particles to real particles at the proper time.
+This feature is accomplished by adding a scan function to monitor the outermost layer of the domain and turn ghost particles to real particles at proper time.\\
 %The computational cost of this additional scan function is low.
-\\
-The data structure, particle and bucket indexing strategies, domain decomposition , dynamic load balancing method and domain adjusting strategies in this paper can be easily adopted by other implementations of any meshfree methods(include SPH).
-The flexibility of data accessing enables implementing of meshless methods for solving of more complicated problems and using of more advanced techniques.\\
+The data structure, particle and bucket indexing strategies, domain decomposition, dynamic load balancing method and domain adjusting strategies in this paper can be easily adopted by other implementations of any meshfree methods(include SPH).
+The flexibility of data accessing enables implementing of meshless methods for solving of more complicated problems and using of more advanced techniques.
 \section{Data Structure and Load Balance}
@@
 \includegraphics[scale=0.34]{Work_flow}
-\caption{Basic work flow for SPH}
+\caption{Basic work flow for SPH code}
 \label{fig:Work_flow}
@@
 Only particles located within support of kernel function will interact.
-Thus, physical properties (position, density, velocity, internal energy, pressure) are updated based on its neighbors.
-So a neighbor search needs to be carried out before updating of physical properties.
+Thus, physical properties (density, velocity, internal energy...) are updated based on its neighbors.
+So a neighbor searching needs to be carried out before updating of physical properties.
 We use buckets which contain all particles associated with a sub-domain
@@
 For example, new particles need to be added at the bottom of the eject vent for JPUE simulation.
-To distinguish particles added at the ``same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
+To distinguish particles added at the "same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
 The time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step.
@@
 \end{itemize}
-We need to emphasize here is that in theory motion of particles will destroy locality established based on initial coordinates of particles.
+We need to emphasize here that in theory motion of particles will destroy locality established based on initial coordinates of particles.
 However, as particles are moving fairly regularly in a JPUE, the locality of most of particles is largely conserved during simulation.
@@
 Information that is contained in particle class can be categorized into six categories: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknowns in governing equations, e.g.
-density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeated computations eg.
-pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases  ) and neighbor information (it is a vector of particle keys in our application).
-Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and owned particles.
+density, velocity), secondary variables (physical properties that can be computed from primitive variables, they are stored to avoid repeated computations eg.
+pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases) and neighbor information (it is a vector of particle keys in our application).
+Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID, affiliate, domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets including its own key) and owned particles.
 Objects defined based on these two classes are then accessed through hash tables.
@@
 But considering the very long conflicts only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash conflicts.
-This decision was made based on numerical experiments.\\
+This decision was made based on numerical experiments.
 \subsubsection{Hash function}
@@
 The SFC sequence passing through centroids of all buckets now becomes a weighted sequence.
-Domain decomposition  is conducted based on the weighted SFC of buckets.\\
+Domain decomposition  is conducted based on the weighted SFC of buckets.
 \begin{table}[t!]
@@
     \hline
-    Particle type & NS & UPME & UPD & UPP &VF &WBC &Total\\
+    Particles & NS & UPME & UPD & UPP &VF &WBC &Total\\
     \hline
@@
     \hline
-    wall ghost & No & No & No & No & No & Yes &0.75\\
-    \hline
-    eruption ghost & No & No & No & Yes & No & No & 0.02\\
-    \hline
-    pressure ghost & No & No & No & No & No & No & 0.00\\
+    wall & No & No & No & No & No & Yes &0.75\\
+    \hline
+    eruption & No & No & No & Yes & No & No & 0.02\\
+    \hline
+    pressure & No & No & No & No & No & No & 0.00\\
     \hline
@@
 \\
-As some of the neighbor particles reside in other
-partitions. A set of  particles and buckets are used to synchronize data across partitions. To minimize communications, data is synchronized only where needed, using non-blocking MPI communications. 
+As some of the neighbor particles reside in other partitions.
+A set of  particles and buckets are used to synchronize data across partitions.
+To minimize communications, data is synchronized only where needed, using non-blocking MPI communications.
 \section{Adjusting of Domain During Simulation}
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.36]{Work_flow_adjust}
+\caption{Work flow that enables domain adjusting feature}
+\label{fig:Work_flow_adjust}
+\end{figure}
 %SPH particles can be viewed as discretization points\cite{diltsg1999moving}. 
@@
 The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such a feature (see figure \ref{fig:Work_flow_adjust}).\\
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.33]{Work_flow_adjust}
-\caption{Work flow that enables domain adjusting feature}
-\label{fig:Work_flow_adjust}
-\end{figure}
 This is implemented by using a involved-flag, in the particle class.
@@
 Other physical properties, like velocity, can also serve as alternative "swich criteria".\\
-\begin{figure*}[!t]
-\centering
-\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.24]{Involved_initial}}
-\label{fig:Involved_initial}
-\hfil
-\subfloat[Computational domain at $t=t0 + \delta t$]{\includegraphics[scale = 0.24]{Involved_domain_adjusting}}
-\label{fig:Involved_domain_adjusting}
-\caption{ Adjusting the domain based on involved-flag of particles}
-\label{fig:Domain_adjusting}
-\end{figure*}
 A similar scheme can also be deployed for buckets resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
@@
 This domain adjusting process is shown in figure \ref{fig:Domain_adjusting}.
-The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}\\
-%
+The work flow with domain adjusting is in figure \ref{fig:Work_flow_adjust}.
+\begin{figure*}[!t]
+\centering
+\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.23]{Involved_initial}}
+\label{fig:Involved_initial}
+\hfil
+\subfloat[Computational domain at $t=t0 + \Delta t$]{\includegraphics[scale = 0.23]{Involved_domain_adjusting}}
+\label{fig:Involved_domain_adjusting}
+\caption{ Adjusting the domain based on involved-flag of particles}
+\label{fig:Domain_adjusting}
+\end{figure*}
 \section{Numerical Test}
-%
-%
 Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
@@
 In current model the initial domain is a 3D box. 
-The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
-At the vent, $T$, $\textbf{v}=\{0,0,150\}^T$, $p=1.01\times10^5 Pa$, $n_{g0}=0.05$ and mass discharge rate $\dot M$ is given. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Velocity is zero for non-slip wall boundary.
-We assume the boundary to be adiabatic and the heat flux is zero on the boundary.
-The pressure of the surrounding atmosphere is specified on pressure boundaries.
-Except for the pressure, density, velocity, and energy will depend on the solution.
+%The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
+At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of solid portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 
+%The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
+Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified on the pressure boundaries. 
+%Except for the pressure, density, velocity, and energy will depend on the solution.
 We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}).
 \subsection{Discretized governing equations with SPH}
-There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, rosswog2009astrophysical, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.
+There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.
 These may be consulted for additional detail.%We will not cover the basic theory of SPH in this paper.
@@
 \begin{equation}
-\textbf{v}_{\alpha b}=\textbf{v}_{\alpha}-\textbf{v}_b
-\end{equation}
-\begin{equation}
-\textbf{v}_{\alpha j}=\textbf{v}_{\alpha}-\textbf{v}_j
+\textbf{v}_{\alpha \beta}=\textbf{v}_{\alpha}-\textbf{v}_{\beta}
 \end{equation}
@@
 \begin{align}
-\dfrac{d \textbf{r}_a}{dt} = \textbf{v} \label{eq:gov-update-pos}
+\dfrac{d \textbf{r}_a}{dt} = \textbf{v}_{\alpha} \label{eq:gov-update-pos}
 \end{align}
@@
 Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
-%
-%
+A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. 
+\begin{figure}[!t]
+\centering
+\subfloat[A plume from Sakurajima Volcano eruption\cite{PlumePhoto}]{\includegraphics[width=3.8cm,height=4.5cm]{plume_photo}}
+\label{fig:plume_photo}
+\hfil
+\subfloat[Contour of mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.5cm]{Plume_simulation}}
+\label{fig:plume_simulation}
+\caption{The volcano plume}
+\label{fig:Plume}
+\end{figure}
 \section{Conclusion}

### 2016-01-08 Paper for submission.tex
cleaned up, ready to submit

@@
  Buffalo, New York 14260\\
-Email: zhixuanc@buffalo.edu, abani@buffalo.edu}
+Email: zhixuanc@buffalo.edu, \\ abani@buffalo.edu}
 \and
 \IEEEauthorblockN{Matthew Jones, Jaroslaw Zola}
-\IEEEauthorblockA{Center for Computational Research\\
+\IEEEauthorblockA{Center for Computational Research, \\
 Department of CSE \\%Computer Science and Engineering
@@
 Crespo\cite{crespo2015dualsphysics} used the GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
-However, most implementation of parallel SPH method presented to date are limited to standard SPH and benchmark problems like dam break, or relatively
-simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume (figure \ref{fig:Plume}) which is essentially a mult-phase turbulent flow,
+However, most implementations of parallel SPH method presented to date are limited to standard SPH and benchmark problems like dam break, or relatively
+simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume (figure \ref{fig:Plume}) which is essentially a multiphase turbulent flow,
 ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction etc. is relatively rare. 
@@
 We define C++ classes which contains all data of particle and bucket.
-As for the management of data, we adopt a hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
+As for the management of data, we adopt hash tables to store pointers to particles and buckets, which give us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
 Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier -- a strategy known to preserve some locality at minimal cost.
@@
 \centering
-\includegraphics[scale=0.34]{Work_flow}
+\includegraphics[scale=0.33]{Work_flow}
 \caption{Basic work flow for SPH code}
@@
 The physical laws (such as conservation laws of mass, momentum and energy) written in the form of partial differential equations need to be transformed into the Lagrangian particle formalism of SPH.
-Using a kernel function that provides the weighted estimate of the field variables in the neighborhood of a discrete point (particle), the integral equations are evaluated as sums over neighbor particles.
+Using a kernel function that provides the weighted estimation of the field variables in the neighborhood of a discrete point (particle), the integral equations are evaluated as sums over neighbor particles.
 Only particles located within support of kernel function will interact.
-Thus, physical properties (density, velocity, internal energy...) are updated based on its neighbors.
+Thus, physical properties (density, velocity, internal energy...) associated to the particle are updated based on its neighbors.
 So a neighbor searching needs to be carried out before updating of physical properties.
 We use buckets which contain all particles associated with a sub-domain
-and are kept fixed in time during the entire simulation, to reduce search cost (since search can now be restricted to only neighboring domains).
-Domain decomposition will be based on an SFC going through centroids of all buckets. A basic work flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
+and are kept fixed in time during the entire simulation, to reduce searching cost (since search can now be restricted to only neighboring domains).
+Domain decomposition will be based on a SFC going through centroids of all buckets. A basic work flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
 Particles need to be added and(or) removed during simulation of JPUE problems.
@@
 The map from coordinates to key is based on SFC.\\
-The SFC \cite{sagan2012space} maps n-dimentional space to a one dimentional sequence. The standard procedure for obtainning SFC is: 
+The SFC \cite{sagan2012space} maps a n-dimentional domain to a one dimentional sequence. The standard procedure for obtainning SFC is: 
 \begin{itemize}
@@
 \item Convert $k_r$ to integer $k$ by multiplying $k_r$ with a very large number and removing decimal part.
-\item All keys are sorted to form a sequence which is SFC. The SFC represents a curve passing through all particles (or centroid of buckets).
+\item All keys are sorted to form a sequence which is SFC. The SFC represents a curve passing through all particles (or centroids of buckets).
 \end{itemize}
@@
 In some situations, new particles need to be added while simulation.
-For example, new particles need to be added at the bottom of the eject vent for JPUE simulation.
+For example, new particles need to be added at the bottom of the eject vent for JPUE simulation (see figure \ref{fig:Particle_adding_with_link}).
 To distinguish particles added at the "same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
-The time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step.
+Then the time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step.
 The map $h_n$ will become:
@@
 pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases) and neighbor information (it is a vector of particle keys in our application).
-Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID, affiliate, domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets including its own key) and owned particles.
+Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID, affiliate, domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets including its own key) and owned particles (it is a vector of particle keys in our application).
 Objects defined based on these two classes are then accessed through hash tables.
@@
 Another option is B-tree.
-We adopt hash table.
+We adopt hash table in our code.
 An implementation of B-tree under a similar situation for mesh based methods can be found in other papers (eg.\cite{patra2003data}).\\
@@
  %We should also do some experiment to optimize our hash table size.
+In addition, Dr.
+Patra's results might no longer reliable due to quick development in CPU architechture, such as changes in the size of caches.
 One way to handle hash conflicts is by using an additional sorted vector attached to the hash table.
@@
 \includegraphics[scale=0.255]{Particle_adding_with_link}
-\caption{Non-uniform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portion of the domain, pointer to these new particles will be stored in external link list.}
+\caption{Non-uniform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles only at a small portion of the whole domain.
+These particles are added at different time steps.
+Pointers to these new particles will be stored in external link lists.}
 \label{fig:Particle_adding_with_link}
@@
 One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as the higher digit and $t$ as the lower digit of the large number.
-However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area.
+However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area (see figure \ref{fig:Domain_adjusting}).
 Only ghost particles for eruption boundary condition will be successively added at the same place: bottom of the vent.
@@
   \centering
-  \caption{Computational Work Load for Each Type of Particle}
+  \caption{Computational Work Load for Different Types of Particles}
   \label{tab:Computational_cost_particles}
@@
     \hline
-    Particles & NS & UPME & UPD & UPP &VF &WBC &Total\\
+    Types & NS & UPME & UPD & UPP &VF &WBC &Total\\
     \hline
@@
 %Domain decomposition will be conducted based the weighted SFC of buckets.
-Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets.
+Figure \ref{fig:SFC_particles_buckets_partition} shows how domain is decomposed based on partition of SFC of buckets.
 The particles are automatically split into several groups along with buckets that contain them.
@@
 As some of the neighbor particles reside in other partitions.
-A set of  particles and buckets are used to synchronize data across partitions.
+A set of "guest" particles and buckets are used to synchronize data across partitions.
 To minimize communications, data is synchronized only where needed, using non-blocking MPI communications.
@@
 \centering
-\includegraphics[scale=0.36]{Work_flow_adjust}
+\includegraphics[scale=0.32]{Work_flow_adjust}
 \caption{Work flow that enables domain adjusting feature}
@@
 We propose a simple strategy to add such feature in our code with low computational cost.
-We add a scan function to monitor the most outside layer of the domain.
+We add a scan function to monitor the outermost layer of the domain.
 When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition.
@@
 The involved particles are particles that have already been affected by the mixing.
-The potentially involved particles are particles that have not been involved in mixing but are adjacent to involved particles and will thus be be involved in the near future.
+The potentially involved particles are particles that have not been involved in mixing but are adjacent to involved particles and will thus be involved in the near future.
 %These not involved particles are particles that far away from the eject source and are still at the initial state.
@@
 So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold ($10^{-5} $ in our simulation).
-Other physical properties, like velocity, can also serve as alternative "swich criteria".\\
+Other physical properties, like velocity, can also serve as alternative "swich criterias".\\
 A similar scheme can also be deployed for buckets resulting in a categorization  of buckets. The additional complexity here is that potentially involved ghost particles must also be accounted for.
@@
 Thus, the test problem used should have such requirements.
-Simulations of a volcanic eruption, which is essentially a muti-phase, turbulent, ejection mixing flow accompanied with microphysics processes will place these demands on the framework.
+Simulations of a volcanic eruption, which is essentially a multi-phase, turbulent, ejection mixing flow accompanied with microphysics processes will place these demands on the framework.
 Thus, we adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem.
@@
 Two types of processors -- 8 core L5630 running at 2.13GHz clock rate with 3GB memory per core on a Mellanox network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic network.
-Each node comprised of two sockets with two of these processors.
+Each node comprised of two sockets with several of these processors.
 Memory and level 3 cache are shared on each node.
@@
 \centering
-\subfloat[A plume from Sakurajima Volcano eruption\cite{PlumePhoto}]{\includegraphics[width=3.8cm,height=4.5cm]{plume_photo}}
+\subfloat[A plume from Sakurajima Volcano eruption\cite{PlumePhoto}]{\includegraphics[width=3.8cm,height=4.4cm]{plume_photo}}
 \label{fig:plume_photo}
 \hfil
-\subfloat[Contour of mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.5cm]{Plume_simulation}}
+\subfloat[Contour of mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}
 \label{fig:plume_simulation}
@@
 Hashtables with external link list are adopted for accessing particles and buckets data.
-Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval.
+Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimized interval.
 The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation.
 Scalability tests on our code shew acceptable strong scalability and good weak scalability.
-3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.
+It was also demonstrated that 3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.
 % An example of a floating figure using the graphicx package.

### 2016-01-08 Paper for submission.tex
small modification

@@
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH. 
-In both SPHysics\cite{dominguez2011optimization} and DualSPHysics\cite{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
+In both SPHysics and DualSPHysics\cite{crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
 This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.

### 2016-01-08 Paper for submission.tex
figure 9

@@
 \and
-\IEEEauthorblockN{Matthew Jones, Jaroslaw Zola}
-\IEEEauthorblockA{Center for Computational Research, \\
-Department of CSE \\%Computer Science and Engineering
+\IEEEauthorblockN{Matthew Jones}
+\IEEEauthorblockA{Center for Computational Research\\%Computer Science and Engineering
 The University at Buffalo\\
 Buffalo, New York 14260\\
-Email: jonesm@buffalo.edu, \\ jzola@buffalo.edu}
-%\and
-%\IEEEauthorblockN{Jaroslaw Zola}
-%\IEEEauthorblockA{The University at Buffalo\\
-%Department of Computer Science and Engineering\\
-% The University at Buffalo\\
-% Buffalo, New York 14260\\
-%Email:jzola@buffalo.edu}
+Email: jonesm@buffalo.edu}
+\and
+\IEEEauthorblockN{Jaroslaw Zola}
+\IEEEauthorblockA{Department of CSE\\
+The University at Buffalo\\
+Buffalo, New York 14260\\
+Email:jzola@buffalo.edu}
 }
@@
 %to distinguish particles added at the same position but different simulation time step. 
-Both particle and background grid are managed by hashtables which ensure quick and flexible accessing, adding and deleting of element.
+Both particle and background grid are managed by hash tables which ensure quick and flexible accessing, adding and deleting of element.
 %Additional link list attached to the hashtable is used to handle hash conflicts.
@@
 \end{equation}
-Where $\textbf{T} \subset [0,\infty)$ is the time step dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.
+Where $\textbf{T} \subset [0,\infty)$ is the time dimension, $\textbf{N}=\lbrace 0, 1, 2, 3...\rbrace$.
 To guarantee locality, sorting of particle keys is majorly based on $k$, that is to say, particle with smaller $k$ always comes before particles with larger $k$.
@@
 \centering
-\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.23]{Involved_initial}}
+\subfloat[Computational domain at $t=t0$]{\includegraphics[scale=0.22]{Involved_initial}}
 \label{fig:Involved_initial}
 \hfil
-\subfloat[Computational domain at $t=t0 + \Delta t$]{\includegraphics[scale = 0.23]{Involved_domain_adjusting}}
+\subfloat[Computational domain at $t=t0 + \Delta t$]{\includegraphics[scale = 0.22]{Involved_domain_adjusting}}
 \label{fig:Involved_domain_adjusting}
@@
 \begin{equation}
-Rm = n_gR_g + n_aR_a  \label{eq:gov-Rm}
+R_m = n_gR_g + n_aR_a  \label{eq:gov-Rm}
 \end{equation}
@@
 %The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
-At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of solid portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 
+At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of gas portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 
 %The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified on the pressure boundaries. 
+Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified, forming pressure boundary conditions. 
 %Except for the pressure, density, velocity, and energy will depend on the solution.
@@
 We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume.
-This will result in a filtered velocity for position updates and additional turbulent terms in momentum and energy equation.
+This will result in a filtered velocity for position updating and additional turbulent terms in momentum and energy equation.
 \subsection{Solver performance}
@@
 Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo.
-Two types of processors -- 8 core L5630 running at 2.13GHz clock rate with 3GB memory per core on a Mellanox network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic network.
+Two types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network.
 Each node comprised of two sockets with several of these processors.
@@
 Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
-A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. 
+A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. Input parameters for this simulation are exactly the same as run P1 in reference \cite{suzuki2005numerical}.
 \begin{figure}[!t]
@@
 \hfil
-\subfloat[Contour of mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}
+\subfloat[Simulation results: mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}
 \label{fig:plume_simulation}
@@
 A time dependent key which is also based on SFC is used as identifier for particle.
-Hashtables with external link list are adopted for accessing particles and buckets data.
+Hash tables with external link list are adopted for accessing particles and buckets data.
 Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimized interval.

### 2016-01-08 Paper for submission.tex
minor modification

@@
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH. 
-In both SPHysics and DualSPHysics\cite{crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
+In both SPHysics and DualSPHysics\cite{crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
 This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
@@
 Storage problems caused by fix-size arrays are thereby also eliminated.
-We define C++ classes which contains all data of particle and bucket.
+We define C++ classes which contain all data of particle and bucket.
 As for the management of data, we adopt hash tables to store pointers to particles and buckets, which give us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
@@
 For example, new particles need to be added at the bottom of the eject vent for JPUE simulation (see figure \ref{fig:Particle_adding_with_link}).
-To distinguish particles added at the "same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into the key.
+To distinguish particles added at the "same place" (the small area, all points within which will be mapped to the same $k$.) at different time steps, we extend the SFC-based key to time-dependent SFC based key by including date of birth of particles into their keys.
 Then the time-dependent SFC based key can be written as: $[k,t]$, where $t$ is the time step.
@@
 pressure, temperature), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases) and neighbor information (it is a vector of particle keys in our application).
-Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID, affiliate, domain information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets including its own key) and owned particles (it is a vector of particle keys in our application).
+Similarly,  Information that is contained in  a bucket class can also be categorized into different categories: ID, affiliate, dimension information (maximum and minimum coordinates, boundary information), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor information (keys of 27 neighbor buckets including its own key) and owned particles (it is a vector of particle keys in our application).
 Objects defined based on these two classes are then accessed through hash tables.
@@
 %Domain decomposition will be conducted based the weighted SFC of buckets.
-Figure \ref{fig:SFC_particles_buckets_partition} shows how domain is decomposed based on partition of SFC of buckets.
+Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets.
 The particles are automatically split into several groups along with buckets that contain them.
@@
 \includegraphics[scale=0.33]{2cases_time}
-\caption{Execution time of test case 1 and test case 2}
+\caption{Wall clock time of test case1 and test case2}
 \label{fig:2cases_time}
@@
 %
-The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively.
+The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively.
 The computational work load of test case 2 is 8 times of that of the test case 1.
@@
 These fluctuations are due to sharing nodes with other users.
-In addition, difference in simulation time for different processors are obvious.
+In addition, difference in simulation time for different processors are obvious.\\
 %That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
@@
 Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
-A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. Input parameters for this simulation are exactly the same as run P1 in reference \cite{suzuki2005numerical}.
+A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. Input parameters for this simulation are exactly the same as "Run P1" in reference \cite{suzuki2005numerical}.
 \begin{figure}[!t]

### 2016-01-08 Paper for submission.tex
last commit

@@
  Buffalo, New York 14260\\
-Email: zhixuanc@buffalo.edu, \\ abani@buffalo.edu}
+Email: zhixuanc@buffalo.edu,\\abani@buffalo.edu}
 \and
@@
 Buffalo, New York 14260\\
-Email:jzola@buffalo.edu}
+Email: jzola@buffalo.edu}
 }

### 2016-01-29 Paper for submission.tex
The final manuscript
01292016

@@
 As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.
-%to distinguish particles added at the same position but different simulation time step. 
 Both particle and background grid are managed by hash tables which ensure quick and flexible accessing, adding and deleting of element.
-%Additional link list attached to the hashtable is used to handle hash conflicts.
 A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance.
@@
 There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.
-These may be consulted for additional detail.%We will not cover the basic theory of SPH in this paper.
+These may be consulted for additional details.
+%We will not cover the basic theory of SPH in this paper.
 The discretized governing equations with SPH are:
@@
 Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
-Two types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network.
-Each node comprised of two sockets with several of these processors.
+Three types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network.
+Each node comprise of two sockets with several of these processors.
 Memory and level 3 cache are shared on each node.
-An important caveat in interpreting these results is that they were obtained in settings where the processors were obtained under normal scheduling (hence readily available) wherein the nodes may be shared with other users.
-Thus our results may suffer from contention for memory, CPU-memory bandwidth and network access.
-As of this writing, we are waiting for exclusive access which will undoubtedly provide better results.
-On the other hand, these results are representative of what will be available during production use.
-%
-The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2, respectively.
+The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2.
 The computational work load of test case 2 is 8 times of that of the test case 1.
@@
 This can be explained by the fact that 2D domain decomposition can get more regular subdomain and as a results, need less communication.
-When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.
+When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.\\
 The weak scalability test is conducted with the same initial domain and various smoothing length.
 Each simulation runs for 400 loops.
-The average number of real particles of each process keep constant at $25900$.
-As shown in figure \ref{fig:weak_loglog}, simulation time are almost constant with some minor fluctutations around the average value.
-These fluctuations are due to sharing nodes with other users.
-In addition, difference in simulation time for different processors are obvious.\\
+The average number of real particles of each process keeps constant at $25900$.
+Tests are carried out under two different settings: the normal scheduling (hence readily available) wherein the nodes may be shared with other users and exclusive access setting.
+Results of normal setting tests may suffer from contention for memory, CPU-memory bandwidth and network access but are representative of what will be available during production use.
+As shown in figure \ref{fig:weak_loglog}, simulation time are almost constant with some minor fluctutations except for an abrupt jump from 132 processors to 620 processors.
+This jump, based on monitor data from nodes health check system on CCR, is due to a daemon which eats up a cpu.
+This inteference happens for all tests whose number of processes larger than or equal to 620.
+The wall clock time of exclusive test is slightly smaller than non-exclusive test which consistent with our anticipation.
+%These fluctuations are due to sharing nodes with other users. In addition, difference in simulation time for different processors are obvious.
+\\
 %That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
@@
 \centering
-\subfloat[A plume from Sakurajima Volcano eruption\cite{PlumePhoto}]{\includegraphics[width=3.8cm,height=4.4cm]{plume_photo}}
+\subfloat[A plume from Sakurajima Volcano eruption\cite{PlumePhoto}]{\includegraphics[width=3.8cm,height=4.0cm]{plume_photo}}
 \label{fig:plume_photo}
 \hfil
-\subfloat[Simulation results: mass fraction of erupted material at 290 seconds]{\includegraphics[width=3.8cm,height=4.4cm]{Plume_simulation}}
+\subfloat[Simulation results: mass fraction of erupted material at 220 seconds]{\includegraphics[width=3.8cm,height=4.0cm]{Plume_simulation}}
 \label{fig:plume_simulation}

### 2016-11-30 Paper for submission.tex
a new branch

@@
 Email: jonesm@buffalo.edu}
-\and
-\IEEEauthorblockN{Jaroslaw Zola}
-\IEEEauthorblockA{Department of CSE\\
-The University at Buffalo\\
-Buffalo, New York 14260\\
-Email: jzola@buffalo.edu}
+%\and
+%\IEEEauthorblockN{Jaroslaw Zola}
+%\IEEEauthorblockA{Department of CSE\\
+%The University at Buffalo\\
+%Buffalo, New York 14260\\
+%Email: jzola@buffalo.edu}
 }
@@
 \begin{abstract}
-This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE).
+This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment (JPUE).
 Background grid is used to reduce neighbors searching cost and to decompose (redecompose) the domain.
-Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid.
-As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.
-Both particle and background grid are managed by hash tables which ensure quick and flexible accessing, adding and deleting of element.
+Space Filing Curve (SFC) based index is adopted to assign an unique identifier to each background grid.
+As simulation of JPUE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.
+Both particle and background grid are managed by hash tables which ensure quick and flexible accessing, adding and deleting of elements.
 A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance.
@@
 Numerical tests show that our code has acceptable strong scalability and good weak scalability.
-These strategies can be further applied to many other implementations of meshfree methods, especially those implementations that require irregular
-particle distributions, adding and deleting of particles and leverage adaptive background grids.
+These strategies can be further applied to many other implementations of meshfree methods, especially those implementations that require irregular particle distributions, adding and deleting of particles and leverage adaptive background grids.
 \end{abstract}
@@
 % no \IEEEPARstart
-SPH is a meshless scheme, initially developed for astrophysical applications by Lucy\cite{lucy1977numerical} and Gingold and Monaghan\cite{gingold1977smoothed}. 
+SPH is a meshless scheme, initially developed for astrophysical applications by Lucy \cite{lucy1977numerical} and Gingold and Monaghan \cite{gingold1977smoothed}. 
 Subsequently it was extended to computational fluid mechanics. 
 A lot of research has appeared on parallellization of SPH in the last few years. 
-Goozee\cite{goozee2003distributed} implemented a simple SPH code using MPI, OpenMP and BSP. 
-Wenbo\cite{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
-Holmes\cite{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
-Dominguez\cite{dominguez2011optimization} presented optimizations for both CPU and GPU parallelization of a SPH method. 
-Ferrari\cite{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme.
+Goozee \cite{goozee2003distributed} implemented a simple SPH code using MPI, OpenMP and BSP. 
+Wenbo \cite{wenbo2014performance} presented a parallel SPH implementation on multi-core CPUs. 
+Holmes \cite{holmes2011framework} presented a simulation framework that enables distributed numerical computing in multi-core shared-memory environments. 
+Dominguez \cite{dominguez2011optimization} presented optimizations for both CPU and GPU parallelization of a SPH method. 
+Ferrari \cite{ferrari2009new} parallelized a 3D SPH code using the message passing interface (MPI) standard, together with a dynamic load balancing strategy to improve the computational efficiency of the scheme.
 Kumar et. al. \cite{kumar2013parallel} implemented a parallel Godunov SPH in simulation of granular flows.
-Crespo\cite{crespo2015dualsphysics} used the GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
+Crespo \cite{crespo2015dualsphysics} used the GPUs to accelerate DualSPHysics by up to two orders of magnitude compared to the performance of the serial version.\\
 However, most implementations of parallel SPH method presented to date are limited to standard SPH and benchmark problems like dam break, or relatively
-simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume (figure \ref{fig:Plume}) which is essentially a multiphase turbulent flow,
+simple scenarios like breaking-waves, floodings etc. Work on more complicated problems, such as eruption of volcano plume (Fig. \ref{fig:Plume}) which is essentially a multiphase turbulent,
 ejection mixing process accompanied by microphysics phenomena like phase change of water, aggregation, reaction etc. is relatively rare. 
@@
 What we need to point out is that the benefit of flexible data management strategies are not only limited to specific implementations of SPH.
-Actually, flexibility in data management are more critical for several advanced techniques of SPH, such as dynamic particle splitting techniques\cite{vacondio2012accurate, feldman2007dynamic}, which will give greater resolution at the area of interest by splitting one large particle to several smaller ones.
+Actually, flexibility in data management are more critical for several advanced techniques of SPH, such as dynamic particle splitting techniques\cite{vacondio2012accurate, feldman2007dynamic}, which will give higher resolution at the area of interest by splitting one large particle to several smaller ones.
 In this paper, we implement SPH to simulate a simplified version of complicated volcanic plume: the JPUE, and develop data management schemes for it.\\
 Among existing CPU parallel SPH schemes, most of them focus on neighbors searching algorithm and dynamic load balancing.
-(eg.\cite{ferrari2009new, crespo2015dualsphysics}).
+(eg.
+\cite{ferrari2009new, crespo2015dualsphysics}).
 Less attention has been paid to developing of more flexible data management schemes for more complicated problems.
@@
 \cite{laszloffy2000simple} for adaptive hp FEM, and \cite{pitman2003computing, patra2005parallel} for FVM).
-Motivated by techniques developed for mesh based methods, we present a complete framwork for parallelizing SPH program with MPI standard model allowing more flexible and efficient data access in this paper.\\
+Motivated by techniques developed for mesh based methods, we present a complete framwork for parallelizing SPH program with MPI standard model allowing more flexible and efficient data access.\\
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many choices possible for this we adopted a background grid which was proposed by Monaghan and Lattanzio\cite{monaghan1985refined} and is quite popular in parallel SPH.
+Of the many possible choices we adopt a background grid which was proposed by Monaghan and Lattanzio\cite{monaghan1985refined} and is quite popular in parallel SPH.
 The background grid is also used for domain decomposition in SPH.
@@
 In both SPHysics and DualSPHysics\cite{crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
-This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
-But adding, deleting and especially accessing of particles are cumbersome.
+This has two advantages: 1) access pattern is more regular and more efficient, 2) it is easy to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
+But adding, deleting and especially accessing of particles are not flexible enough.
 Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
 Storage problems caused by fix-size arrays are thereby also eliminated.
-We define C++ classes which contain all data of particle and bucket.
-As for the management of data, we adopt hash tables to store pointers to particles and buckets, which give us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
+We define C++ classes which contain all data for particle and bucket.
+As for the management of data, we adopt hash tables to store pointers to objects of particles and buckets, which give us not only flexibility of deleting and adding element, but also quicker access compared with linked list.
 Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier -- a strategy known to preserve some locality at minimal cost.
@@
 To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation needs.
-For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluid.
+For JPUE simulation, such feature will greatly reduce computational cost by avoiding computing of uninfluenced fluids.
 This feature is accomplished by adding a scan function to monitor the outermost layer of the domain and turn ghost particles to real particles at proper time.\\
 %The computational cost of this additional scan function is low.
-The data structure, particle and bucket indexing strategies, domain decomposition, dynamic load balancing method and domain adjusting strategies in this paper can be easily adopted by other implementations of any meshfree methods(include SPH).
+The data structure, particle and bucket indexing strategies, domain decomposition, dynamic load balancing method and domain adjusting strategies in this paper can be easily adopted by other implementations of any meshfree methods (include SPH).
 The flexibility of data accessing enables implementing of meshless methods for solving of more complicated problems and using of more advanced techniques.
@@
 and are kept fixed in time during the entire simulation, to reduce searching cost (since search can now be restricted to only neighboring domains).
-Domain decomposition will be based on a SFC going through centroids of all buckets. A basic work flow of our SPH code is shown in figure \ref{fig:Work_flow}.\\
-Particles need to be added and(or) removed during simulation of JPUE problems.
-Wall boundary conditions in SPH can be imposed either by adding a force term or using ghost particles.
+A basic work flow of our SPH code is shown in figure \ref{fig:Work_flow}.
+Domain is decomposed based on a SFC going through centroids of all buckets. Our data management strategies will also guarantee flexible adding, deleting and accessing of data.
+%Wall boundary conditions in SPH can be imposed either by adding a force term or using ghost particles.
 We adopt the latter one in our simulation.
@@
 New particles also need to be added for the eruption boundary condition.
-We will describe in this section strategies which satisfy these requirements.
+%We will describe in this section strategies which satisfy these requirements.
 \subsection{SFC based indexing}
@@
 \end{itemize}
-We need to emphasize here that in theory motion of particles will destroy locality established based on initial coordinates of particles.
+We need to emphasize here that, in theory, motion of particles will destroy locality established based on initial coordinates of particles.
 However, as particles are moving fairly regularly in a JPUE, the locality of most of particles is largely conserved during simulation.
@@
 Patra's results might no longer reliable due to quick development in CPU architechture, such as changes in the size of caches.
-One way to handle hash conflicts is by using an additional sorted vector attached to the hash table.
+One way to handle hash conflicts is using an additional sorted vector attached to the hash table.
 When several keys hash to the same slot, a vector will be created.
@@
 Choosing of proper way to handle hash conflicts greatly depends on the problem itself.
-For the test problem in this paper, successively adding particles at the bottom of the eject vent will lead to hash conflicts of many particles, which implies a long link list.
-But considering the very long conflicts only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), we still choose the link list to handle hash conflicts.
-This decision was made based on numerical experiments.
+For the test problem in this paper, successively adding particles at the bottom of the eject vent will lead to hash conflicts of many particles, which will lead to a long link list.
+But considering the conflicts only occur on several slots among millions (see figure \ref{fig:Particle_adding_with_link}), the link list is a better choice for us.
+%This decision was made based on numerical experiments.
 \subsubsection{Hash function}
@@
 \subsection{Load balancing strategy}
-\subsubsection{Weighted work load}
+\subsubsection{Weighted work load} \label{sec:Weighted_work_load}
 Particles used in the test problem can be categorized into four types based on the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle.
 Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Domain_adjusting}).
-As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data.
-Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket.
+As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles calibrated by profilling data.
+Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weights of all particles within the bucket.
 The SFC sequence passing through centroids of all buckets now becomes a weighted sequence.
@@
     %\hhline{|=|=|=|=|=|=|=|=|}
-    Real & Yes & Yes & Yes & Yes & Yes & No & 2.00\\
+    Real & Yes & Yes & Yes & Yes & Yes & No & 1.98\\
     \hline
@@
 As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition.
-A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Kumar et al \cite{kumar2013parallel}.
+%A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Kumar et al \cite{kumar2013parallel}.
 A comparison of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger.
@@
 \end{figure} 
-%SPH particles can be viewed as discretization points\cite{diltsg1999moving}. 
 As a Lagrangian method, SPH is able to automatically define and adjust its computational domain as the position of the discretization points are updated at every time step.  The 
@@
 However, for JPUE simulation (and many similar/related phenomena), where some fluid ejects into stationary fluid and gets mixed, the domain-adjusting feature of SPH cannot be taken advantage of as all of the stationary fluid must be represented.
-%Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation.
-A lot of CPU time will be spent on computing associated with these stationary particles.
-If simulating of stationary particles can be avoided, the computational cost will be reduced greatly.
-%To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation.
+A lot of CPU time will be spent on computing associated with these stationary particles.  If simulating of stationary particles can be avoided, the computational cost will be reduced greatly.
 We propose a simple strategy to add such feature in our code with low computational cost.
 We add a scan function to monitor the outermost layer of the domain.
-When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition.
+When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for boundary conditions.
 The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such a feature (see figure \ref{fig:Work_flow_adjust}).\\
@@
 The potentially involved particles are particles that have not been involved in mixing but are adjacent to involved particles and will thus be involved in the near future.
-%These not involved particles are particles that far away from the eject source and are still at the initial state.
-For these not involved particles, it is meaningless to update its physical properties.
-As a consequence, searching for neighbors and data communitation for these particles are also unnecessary.
 All communication and computation associated with uninvolved particles can be ignored. That is to say, only potential involved and involved particles need to be simulated.
@@
 \end{figure*}
+
 \section{Numerical Test}
-Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
-Thus, the test problem used should have such requirements.
-Simulations of a volcanic eruption, which is essentially a multi-phase, turbulent, ejection mixing flow accompanied with microphysics processes will place these demands on the framework.
-Thus, we adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem.
-In this model, one phase is air while another phase is ejected material.
-\subsection{Governing equations\cite{suzuki2005numerical} and boundary conditions}
-Based on Navier-Stokes equations and several simplifications, the governing equations in Eulerian form are:
-%the following assumptions are made:
-%\begin{itemize}
-%\item Molecular Viscosity is neglected since eddy viscosity due to turbulence is dominant.
-As heat conduction is much smaller than turbulent heat exchange \cite{oberhuber1998volcanic}, We only consider turbulent heat exchange.
-%\item Erupted material consist of solid with different size and mixture of gas (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2).
-Air (also a mixture of different gas) is assumed to be another phase (phase 1).
-%\item We assume immediate thermodynamic equilibrium so that no separate energy equation is needed for each phase.
-As a result, there is only one energy equation for both phases and heat transfer effect between different phases will be ignored.
-%\item We assume immediate dynamic equilibrium so that no separate momentum equation is needed for each phase.
-As a result, there is only one vector moment equation for both phases.
-Drag force term will not show up in the momentum equation as two phases always move with the same velocity.
-%\item Because of above assumptions, all other micro-physics process (like phase change of $H_2O$, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model.
-%\end{itemize}
-%Finally the governing equations for this model in Eulerian form are:
-\begin{eqnarray} %%I did not notice that I can add equation like this, useful, haha
-\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
-%\end{equation}
-%\begin{equation}
-\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) &=& 0 \label{eq:gov-cs-ks}\\
-%\end{equation}
-%\begin{equation}
-\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) &=& \rho \textbf{g} \label{eq:gov-cs-v} \\
-%\end{equation}
-%\begin{equation}
-\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] &=& \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
-\end{eqnarray}
-$\xi$ is the mass fraction of ejected material.
-$E = e + K $ is total energy which is summation of kinetic energy $K$ and internal energy $e$.
-An additional equation is required to close the system. In this model, the equation for closing the system is an EOS:
-\begin{align}
-p = (\gamma_m - 1)\rho e \label{eq:EOS}
-\end{align}
-Where 
-\begin{equation}
-\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
-\end{equation}
-\begin{equation}
-R_m = n_gR_g + n_aR_a  \label{eq:gov-Rm}
-\end{equation}
-\begin{equation}
-C_{vm} = n_s C_{vs} + n_g C_{vg} + n_a C_{va} \label{eq:gov-Cvm}
-\end{equation}
-\begin{equation}
-n_a = 1 - \xi \label{eq:gov-na}
-\end{equation}
-\begin{equation}
-n_g = \xi n_{g0} \label{eq:gov-ng}
-\end{equation}
-\begin{equation}
-n_s = \xi - n_g \label{eq:gov-ns}
-\end{equation}
-Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscript 
-$m$ represents mixture of ejected material and air, $s$ is solid portion in ejected material, $g$ is gas portion in the ejected material and $a$ is air.\\
-In current model the initial domain is a 3D box. 
-%The boundaries are categorized into eruption vent (a circle area at the center of the bottom), wall boundary (box bottom), pressure boundary (Other faces of the box).
-At the eruption vent, temperature $T$, velocity $\textbf{v}$, pressure $p$, initial mass fraction of gas portition $n_{g0}$ and mass discharge rate $\dot M$ is given. 
-%The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Non-slip, adiabatic wall boundary condition is imposed on the wall. The pressure of the surrounding atmosphere is specified, forming pressure boundary conditions. 
-%Except for the pressure, density, velocity, and energy will depend on the solution.
-We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}).
-\subsection{Discretized governing equations with SPH}
-There are several review papers \cite{monaghan1992smoothed, monaghan2005smoothed, price2012smoothed, monaghan2012smoothed} which provide a pretty comprehensive descriptions of SPH.
-These may be consulted for additional details.
-%We will not cover the basic theory of SPH in this paper.
-The discretized governing equations with SPH are:
-\begin{equation}
-<\rho_a^a>=\sum m_b w_{ab} (h_a) \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_i^{sg}>=\sum_j m_j w_{ij} (h_i) \label{eq:gov-sph-d2}
-\end{equation}
-\begin{equation}
-\begin{split}
-<\dfrac{d \textbf{v}_{\alpha}}{d t}>= &
--\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}) \\ &\bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
--\sum_j [m_j (\dfrac{p_j}{\rho_j^2} \\ & + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
-+\textbf{g}
-\end{split} 
-\label{eq:gov-sph-v}
-\end{equation}
-\begin{equation}
-\begin{split}
-<\dfrac{d e_{\alpha}}{d t}>=&
- 0.5\sum_b [m_b \textbf{v}_{\alpha b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b})\\ & \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
-+0.5\sum_j [m_j \textbf{v}_{\alpha b}(\dfrac{p_j}{\rho_j^2} \\ & + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
-\end{split} 
-\label{eq:gov-sph-e}
-\end{equation}
-Where
-$\rho_a^a$ is density of phase 1 (air).
-$\rho_i^{sg}$ is density of phase 2 (erupted material).
-$\rho=\rho^a + \rho^{sg}$ is density of mixture of phase 1 and phase 2.
-\begin{equation}
-\textbf{v}_{\alpha \beta}=\textbf{v}_{\alpha}-\textbf{v}_{\beta}
-\end{equation}
-$w_{ab} (h_a)= w(\textbf{r}_a-\textbf{r}_b,h_a)$ is smoothing kernel.
-%$\bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})$-->need double check
-$\Pi$ is artificial viscosity term \cite{monaghan1992smoothed}.
-Index $a$, $b$ is for phase 1.
-Index $i$, $j$ is for phase 2.
-Index $\alpha$, $\beta$ can be index of either phase 1 or phase 2.
-The position of each particle is updated according to the following equation.
-\begin{align}
-\dfrac{d \textbf{r}_a}{dt} = \textbf{v}_{\alpha} \label{eq:gov-update-pos}
-\end{align}
-Free surface flows are inherently turbulent.
-We adopt $SPH-\varepsilon$ method developed by Monaghan\cite{monaghan2011turbulence} to capture turbulence in the plume.
-This will result in a filtered velocity for position updating and additional turbulent terms in momentum and energy equation.
-\subsection{Solver performance}
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.33]{2cases_time}
-\caption{Wall clock time of test case1 and test case2}
-\label{fig:2cases_time}
-\end{figure}
-%
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.33]{2cases_efficiency}
-\caption{Influence of total work load on strong scalability}
-\label{fig:2cases_efficiency}
-\end{figure}
-%
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.33]{2D_vs_3D_efficiency}
-\caption{Strong scalability of 3D domain decomposition and 2D domain decomposition}
-\label{fig:2D_vs_3D_efficiency}
-\end{figure}
-%
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{weak_loglog}
-\caption{Weak scalability test results}
-\label{fig:weak_loglog}
-\end{figure}
-Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
-Three types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network.
-Each node comprise of two sockets with several of these processors.
-Memory and level 3 cache are shared on each node.
-The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2.
-The computational work load of test case 2 is 8 times of that of the test case 1.
-The simulations run for 20s physical time.
-Wall clock time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}.
-Linear speed up is observed when number of processes is smaller than 16.
-Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load.
-We also compared the performance of 2D domain decomposition and 3D domain decomposition.
-As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processes.
-This can be explained by the fact that 2D domain decomposition can get more regular subdomain and as a results, need less communication.
-When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.\\
-The weak scalability test is conducted with the same initial domain and various smoothing length.
-Each simulation runs for 400 loops.
-The average number of real particles of each process keeps constant at $25900$.
-Tests are carried out under two different settings: the normal scheduling (hence readily available) wherein the nodes may be shared with other users and exclusive access setting.
-Results of normal setting tests may suffer from contention for memory, CPU-memory bandwidth and network access but are representative of what will be available during production use.
-As shown in figure \ref{fig:weak_loglog}, simulation time are almost constant with some minor fluctutations except for an abrupt jump from 132 processors to 620 processors.
-This jump, based on monitor data from nodes health check system on CCR, is due to a daemon which eats up a cpu.
-This inteference happens for all tests whose number of processes larger than or equal to 620.
-The wall clock time of exclusive test is slightly smaller than non-exclusive test which consistent with our anticipation.
-%These fluctuations are due to sharing nodes with other users. In addition, difference in simulation time for different processors are obvious.
-\\
-%That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
-%When we request 128 processes, 64 processes will be totally idle at the beginning.
-As simulation goes, the domain will expand and more and more idle processes will start getting tasks.
-That's why we still get some speed up when number of process is 128.
-3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.
-%
-%
-For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption.
-When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$.
-This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
-Numercial test shows that simulation time of the test problem is reduce to $\frac{1}{4}$ of original simulation time when we adopt the domain adjusting strategy in our code.
-A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}. Input parameters for this simulation are exactly the same as "Run P1" in reference \cite{suzuki2005numerical}.
 \begin{figure}[!t]
@@
 \end{figure}
+Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
+Thus, the test problem used should have such requirements.
+Simulations of a volcanic eruption, will place these demands on the framework.
+Thus, we adopt a two phase volcanic plume model\cite{suzuki2005numerical} as our test problem.
+Governing equation, boundary condition and input parameters can be found in \cite{suzuki2005numerical}.
+Details about discretization of the governing equations will not be covered in this paper, neither.
+A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}.
+Input parameters for this simulation are exactly the same as "Run P1" in reference \cite{suzuki2005numerical}.
+%
+\subsection{Scalability test}
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.33]{strong_scale}
+\caption{Influence of total work load on strong scalability.
+It will exceed the time limit on CCR to do a serial simulation for case 2, so the serial simulation time is estimated by assuming that speed up of case 1 and case 2 are the same at 64 processors.}
+\label{fig:2cases_efficiency}
+\end{figure}
+%
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.35]{weak_scale}
+\caption{Weak scalability test results}
+\label{fig:weak_loglog}
+\end{figure}
+Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
+Three types of processors -- 8 core L5520/L5630 running at 2.27/2.13GHz clock rate with 3GB memory per core on a Mellanox Infiniband network and 12 core E5645 running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband network.
+Each node comprise of two sockets with several of these processors.
+Memory and level 3 cache are shared on each node.
+The simulational domain is a box with initial ddimension $[-4.8km, 4.8km]\times [-4.8km, 4.8km] \times [0km, 6km]$ for case 1 and $[-10km, 10km]\times [-10km, 10km] \times [0km, 5km]$ for case 2.
+The smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case 1 and test case 2.
+So the computational work load of test case 2 is larger than that of the test case 1.
+The simulations run for 20s physical time.
+Parallel speed up are shown in Fig.
+\ref{fig:2cases_efficiency}.
+Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load.
+The weak scalability test is conducted with the same initial domain as test case 1 and various smoothing length.
+Each simulation runs for 400 time steps.
+The average number of real particles of each process keeps constant at $25900$, so the average work load for each processor are keep constant.
+As shown in figure \ref{fig:weak_loglog}, simulation time are almost constant with some minor fluctutations.\\
+%
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[scale=0.33]{2D_vs_3D_efficiency}
+%\caption{Strong scalability of 3D domain decomposition and 2D domain decomposition}
+%\label{fig:2D_vs_3D_efficiency}
+%\end{figure}
+%
+% We also compared the performance of 2D domain decomposition and 3D domain decomposition.
+As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processes.
+This can be explained by the fact that 2D domain decomposition can get more regular subdomain and as a results, need less communication.
+When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.\\
+\subsection{Effect of workload check interval and calibrated particle weight}
+\begin{table}[t!]
+  \renewcommand{\arraystretch}{1.2}
+  \centering
+  \caption{Simulation Time for Same Particle Weight and Different Particle Weight }
+  \label{tab:same_diff_particle_weight}
+  \begin{tabular}{|c|r|r|r|r|}
+    \hline
+    Physics time & 10 s & 20ss & 30 s & 40 s \\
+    \hline
+    %\hhline{|=|=|=|=|=|}
+    Same weight & 1141.7 & 4119.4 & 10371.0 & 12453.7\\
+    \hline
+    Different weight & 1108.2 & 4057.0 & 10281.5 & 12166.3\\
+    \hline
+  \end{tabular}
+\end{table}
+In section \ref{sec:Weighted_work_load}, we proposed to use different particle weight for different types of particles when estimate the weight of buckets.
+The effect of using different particle weight is demonstrated in table \ref{tab:same_diff_particle_weight}.
+The simulation time is reduced by using different particle weights.
+However, the effect is not as significant as our expectation.
+One possible explanation is that real particles are the major population, so the load imbalance caused by assignning improper weight value to ghost particles is small.\\
+The best load balance check interval is calibrated based on a series of simulations with different load balance scan/repartition intervals.
+From 0 - 50 seconds, the interval of 1 second shows a better load balance than interval of 2 second.
+However, for the simulation of 50 - 100 seconds, interval of 2 second is better.
+This implies that lose of load balance is faster and requires a more  frequent re-decomposition of domain at the initial stage of simulation.
+This is consistent with the plume development process: the domain grows quickly at beginning as erupted material ejects into the environment and spreads quickly.
+Afterwards, the spreading speed of the front edge slows down due to viscosity and momentum exchange and domain growth slows down, too.\\
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.35]{int_bar}
+\caption{The influence of different load balance check intervals on simulation time. The bar values are the extra simulation time in log scale. The optimized interval is 3s.}
+\label{fig:check_int}
+\end{figure}
+To figure out the influence of hash table size on performance, a series of simulations up tp 100 seconds of physics time are conducted.
+Only background buckets hash table size is varied.
+The results are shown in Fig.
+\ref{fig:bg_hashtb_size}.
+Based on these four simulations, the best hash table size should be 600 in a average sense.
+However, at the first period of simulation, 0 - 10 seconds, hash table size of 100 performs better than 600.
+Laterly as number of involved buckets increase due to domain expanding, a larger hash table size performs better.
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.35]{Bg_table_size}
+\caption{The influence of different bucket hash table size on simulation time.
+The bar values are calculated by deducting the minimum simulation time among all simulations from simulation time.
+The y axis is also in log scale.}
+\label{fig:bg_hashtb_size}
+\end{figure}
+We need to emphasize that, these optimized parameters, including weight of particles, load blance checking interval and hashtable size, are case sensitive and need to be re-calibrated for different applications and hardware architecture.
+%
+\subsection{Domain adjusting}
+\begin{table}[t!]
+  \renewcommand{\arraystretch}{1.2}
+  \centering
+  \caption{Computational Work Load of Extra Steps for Domain Adjusting}
+  \label{tab:Computational_cost_doamin_adj}
+  \begin{tabular}{|c|r|r|r|r|r|r|}
+    \hline
+    Cost & UPME & UPP & ADPP & ADWP & SWCH & SCN\\
+    \hline
+    %\hhline{|=|=|=|=|=|=|=|}
+    Total time (s) & 1088.97 & 36.11 & 13.02 & 7.14 & 3.87 & 3.20 \\
+    \hline
+    Called times & 101 & 101 & 3 & 3 & 3 & 101 \\
+    \hline
+  \end{tabular}
+\end{table}
+For the test problem in this paper, the volcanic plume will finally reach to a region of $[-10km \,\,\, 10km] \times [-10km\,\,\,10km] \times [0km\,\,\,20km]$ after around 300 seconds of eruption.
+When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-3km\,\,\,3km] \times [-3km\,\,\,3km] \times [0km\,\,\,6km]$.
+This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
+On the other hand, additional functions (see Fig.
+\ref{fig:Work_flow_adjust}) for domain adjusting will add extra computational cost.
+Table \ref{tab:Computational_cost_doamin_adj} shows the extra computational cost corresponding to these functions.
+In table \ref{tab:Computational_cost_doamin_adj}, SWCH is short for step that switch pressure ghost particle to real particle, ADPP is short for adding new pressure ghost particles, ADWP is short for adding wall ghost particles, SCN is short for scanning the most outside layer of the domain.
+Profilling data for momentum and energy update (UPME) and position update(UPP) is also included in the table for the purpose of comparison.
+As we can see from the table, the cost of SCN is neglectable compared with UPP fucntion, which is the cheapest function as shown in table \ref{tab:Computational_cost_steps}.
+Other three functions, ADPP, ADWP and SWCH, only called fewer times during the simulation, and as a consequence, the extra computational cost due to them are also neglectable.
+To summarize, the additional cost caused by domain adjusting functions is ignorable.
+\begin{figure}[!t]
+\centering
+\includegraphics[scale=0.35]{adj_vs_no}
+\caption{The effect of domain adjusting on simulation time.
+The figure on the top shows simulation time without domain adjusting, the figure on the bottom shows simulation time with domain adjusting.
+Different bins represent simulation time up to specific physics time indicated by $x$ axis.}
+\label{fig:adj_vs_no}
+\end{figure}
+Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy in our code.
 \section{Conclusion}

### 2016-12-07 Paper for submission.tex
a small modification
updated domain adjusting data

@@
     \hline
-    update position & 0.02 & UPP\\
+    update position & 0.01 & UPP\\
     \hline
@@
     %\hhline{|=|=|=|=|=|=|=|=|}
-    Real & Yes & Yes & Yes & Yes & Yes & No & 1.98\\
+    Real & Yes & Yes & Yes & Yes & Yes & No & 1.97\\
     \hline
@@
 \end{figure}
-Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition.
-Thus, the test problem used should have such requirements.
+Our target here is to develop data management and parallelization strategies for more complicated implementations of SPH which demand quick and flexible data access, deletion and addition of particles.
+Thus, the test problem should have such requirements.
 Simulations of a volcanic eruption, will place these demands on the framework.
@@
 Details about discretization of the governing equations will not be covered in this paper, neither.
-A typical simulation result of volcano plume is shown in figure \ref{fig:Plume}.
+A typical simulation result of volcano plume by our code is shown in figure \ref{fig:Plume}.
 Input parameters for this simulation are exactly the same as "Run P1" in reference \cite{suzuki2005numerical}.
@@
 When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128.\\
-\subsection{Effect of workload check interval and calibrated particle weight}
+\subsection{Effect of workload check interval, hash table size and calibrated particle weight}
 \begin{table}[t!]
@@
     %\hhline{|=|=|=|=|=|=|=|}
-    Total time (s) & 1088.97 & 36.11 & 13.02 & 7.14 & 3.87 & 3.20 \\
+    Total time (s) & 2954.8 & 38.55 & 21.51 & 8.88 & 0.08 & 7.72 \\
     \hline
-    Called times & 101 & 101 & 3 & 3 & 3 & 101 \\
+    Called times & 201 & 201 & 3 & 3 & 2 & 201 \\
     \hline
@@
 In table \ref{tab:Computational_cost_doamin_adj}, SWCH is short for step that switch pressure ghost particle to real particle, ADPP is short for adding new pressure ghost particles, ADWP is short for adding wall ghost particles, SCN is short for scanning the most outside layer of the domain.
-Profilling data for momentum and energy update (UPME) and position update(UPP) is also included in the table for the purpose of comparison.
-As we can see from the table, the cost of SCN is neglectable compared with UPP fucntion, which is the cheapest function as shown in table \ref{tab:Computational_cost_steps}.
-Other three functions, ADPP, ADWP and SWCH, only called fewer times during the simulation, and as a consequence, the extra computational cost due to them are also neglectable.
+Profilling data for momentum and energy update (UPME) and position update (UPP) is also included in the table for the purpose of comparison.
+As we can see from the table, the cost of SCN is even smaller than UPP fucntion, which is the cheapest function (as shown in table \ref{tab:Computational_cost_steps}) in the regular framework (see Fig \ref{fig:Work_flow}).
+Other three functions, ADPP, ADWP and SWCH, only called fewer times during the simulation, and as a consequence, the extra computational cost due to them are neglectable.
 To summarize, the additional cost caused by domain adjusting functions is ignorable.

### 2016-12-22 Paper for submission.tex
updated the most recent strong scalability test

@@
 \subsection{Scalability test}
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[scale=0.33]{strong_scale}
+%\caption{Influence of total work load on strong scalability.
+It will exceed the time limit on CCR to do a serial simulation for case 2, so the serial simulation time is estimated by assuming that speed up of case 1 and case 2 are the same at 64 processors.}
+%\label{fig:2cases_efficiency}
+%\end{figure}
 \begin{figure}[!t]
 \centering
-\includegraphics[scale=0.33]{strong_scale}
-\caption{Influence of total work load on strong scalability.
-It will exceed the time limit on CCR to do a serial simulation for case 2, so the serial simulation time is estimated by assuming that speed up of case 1 and case 2 are the same at 64 processors.}
+\subfloat[a]{\includegraphics[scale=0.33]{strong_scale}}
+\label{fig:plume_photo}
+\hfil
+\subfloat[b. zoomed view of a]{\includegraphics[scale=0.33]{strong_scale_zoom}}
+\label{fig:plume_simulation}
+\caption{Influence of total work load on strong scalability. Fig. b is the zoomed view of Fig a. It is obviously shown that strong scalability is better when problem size is larger.}
 \label{fig:2cases_efficiency}
